# Silvicultural Metrics{#silv}

```{r udp-silv, include=FALSE, eval=TRUE, echo=FALSE}
##################################################################################################
## USER DEFINED PARAMETERS
##################################################################################################
### Only include trees above a specific DBH
  # e.g., set to match the minimum tree DBH for field sampling of overstory trees
  # !!!!!!!!!!!!!!!!!!!!!!INCLUSIVE ... i.e. ostory_dbh >= ostory_min_tree_dbh_cm
ostory_min_tree_dbh_cm = 5*2.54 # 3 in = 7.62 cm

### Only include trees under a specific DBH
  # e.g., set to match the minimum tree DBH for field sampling of overstory trees
  # !!!!!!!!!!!!!!!!!!!!!!EXCLUSIVE ... i.e. ustory_dbh < ustory_max_tree_dbh_cm
ustory_max_tree_dbh_cm = 5*2.54 # 3 in = 7.62 cm
##################################################################################################
## USER DEFINED PARAMETERS
##################################################################################################
```

Overstory metrics are calculated using the UAS SfM-derived DBH and height measurements, where DBH is estimated using a [local DBH-Height allometric relationship](#local_mod). [Validation](#ostory_valid) of the UAS SfM-derived DBH and height measurements was completed using field-based sampling data. In the following analysis and reporting, overstory includes any tree where the DBH is `r scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)` in. (`r scales::comma(ostory_min_tree_dbh_cm,accuracy = 0.01)` cm) or greater.

Understory (i.e. regeneration) metrics are calculated using the field-based sampling data reviewed in [this section](#ustory_valid). The following analysis and reporting, understory includes any tree where the DBH is less than `r scales::comma(ustory_max_tree_dbh_cm/2.54,accuracy = 0.1)` in. (`r scales::comma(ustory_max_tree_dbh_cm,accuracy = 0.01)` cm) or greater. 

```{r silv-calc, results='hide'}
# Common silvicultural metrics are calculated for the entire extent.
  # Note, that stand-level summaries can be computed if stand vector data is provided.
  # metrics include:
    # "n_trees"
    # "stand_area_ha"
    # "trees_per_ha"
    # "mean_dbh_cm"
    # "qmd_cm"
    # "mean_tree_height_m"
    # "loreys_height_m"
    # "basal_area_m2"
    # "basal_area_m2_per_ha"

### stand-level summaries
  silv_metrics = harvests_trees %>%
    sf::st_drop_geometry() %>%
    dplyr::filter(dbh_cm >= ostory_min_tree_dbh_cm) %>% 
    dplyr::ungroup() %>%
    dplyr::group_by(suid,stand_area_ha) %>%
    dplyr::summarise(
      n_trees = dplyr::n_distinct(treeID)
      , mean_dbh_cm = mean(dbh_cm, na.rm = T)
      , mean_tree_height_m = mean(tree_height_m, na.rm = T)
      , loreys_height_m = sum(basal_area_m2*tree_height_m, na.rm = T) / sum(basal_area_m2, na.rm = T)
      , basal_area_m2 = sum(basal_area_m2, na.rm = T)
      , sum_dbh_cm_sq = sum(dbh_cm^2, na.rm = T)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      trees_per_ha = (n_trees/stand_area_ha)
      , basal_area_m2_per_ha = (basal_area_m2/stand_area_ha)
      , qmd_cm = sqrt(sum_dbh_cm_sq/n_trees)
    ) %>%
    dplyr::select(-c(sum_dbh_cm_sq)) %>%
    # convert to imperial units
    calc_imperial_units_fn() %>%
    dplyr::select(
      "suid"
      , "n_trees"
      , "stand_area_ha"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2"
      , "basal_area_m2_per_ha"
      # imperial
      , "stand_area_ac"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2"
      , "basal_area_ft2_per_ac"
    )

### export tabular
  write.csv(
      silv_metrics
      , paste0(delivery_dir, "/stand_overstory_silv_metrics.csv")
      , row.names = F
    )

  # join with spatial data
  silv_metrics = harvests %>% 
    # join with plot data data
    dplyr::inner_join(
      silv_metrics
      , by = dplyr::join_by("suid")
    )

######################################################
# UNDERSTORY SUMMARY
######################################################
ustory_silv_metrics = harvests_regen_plt_diam %>% 
  sf::st_drop_geometry() %>%
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("diam_inch_class")
      , .fn = as.numeric
    )
  ) %>% 
  dplyr::filter(
    diam_inch_class_lower < ustory_max_tree_dbh_cm
  ) %>% 
  dplyr::group_by(suid, regen_size_class_diam) %>% 
  dplyr::summarise(
    number_of_plots = dplyr::n_distinct(plot_dir_id)
    , max_diam_inch_class_upper = max(diam_inch_class_upper)
    , sum_num_trees = sum(number_of_trees_in_size_class)
    , sampled_area_ac = sum(plot_area_ac)
    , sampled_area_ha = sum(plot_area_ha)
  ) %>% 
  dplyr::mutate(
    size_class_trees_per_ac = sum_num_trees/sampled_area_ac
    , size_class_trees_per_ha = sum_num_trees/sampled_area_ha
  ) %>% 
  dplyr::group_by(suid) %>% 
  dplyr::mutate(
    unit_trees_per_ac = sum(size_class_trees_per_ac)
    , unit_trees_per_ha = sum(size_class_trees_per_ha)
  ) %>% 
  dplyr::ungroup()
  
### export tabular
  write.csv(
      ustory_silv_metrics
      , paste0(delivery_dir, "/stand_understory_silv_metrics.csv")
      , row.names = F
    )
```

## Overstory Summary

### Distribution of silvicultural metrics

Distribution of silvicultural metrics across the `r nrow(silv_metrics)` stands in this analysis.

```{r site-metrics}
silv_metrics %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    "suid"
    , "stand_area_ac"
    , "n_trees"
    , "trees_per_ac"
    , "mean_dbh_in"
    , "qmd_in"
    , "mean_tree_height_ft"
    , "loreys_height_ft"
    , "basal_area_ft2_per_ac"
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(suid), names_to = "metric", values_to = "val"
  ) %>% 
  # order metrics
  dplyr::mutate(
    metric = factor(
      metric
      , ordered = T
      , levels = c(
          "stand_area_ac"
          , "n_trees"
          , "trees_per_ac"
          , "mean_tree_height_ft"
          , "loreys_height_ft"
          , "mean_dbh_in"
          , "qmd_in"
          , "basal_area_ft2_per_ac"
        )
      , labels = c(
          latex2exp::TeX("Stand Area (ac)", output = "character")
          , latex2exp::TeX("Number of Trees", output = "character")
          , latex2exp::TeX("Trees $ac^{-1}$")
          , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
          , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
          , latex2exp::TeX("Mean DBH (in)", output = "character")
          , latex2exp::TeX("QMD (in)", output = "character")
          , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
        )
    )
  ) %>% 
  ggplot(
    mapping = aes(
      x = val
      , y = 1
    )
  ) +
  geom_boxplot(fill = "gray", width = 0.7, outlier.shape = NA) +
  geom_point(mapping = aes(group = suid)) +
  facet_wrap(facets = vars(metric), ncol = 2, scales = "free_x", labeller = label_parsed) +
  scale_x_continuous(breaks = scales::extended_breaks(6)) +
  scale_y_continuous(limits = c(0.5,1.5)) +
  labs(
    y = ""
    , x = ""
    , title = paste0(
        "Overstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) Distribution of Silvicultural Metrics"
      )
    , subtitle = paste(nrow(silv_metrics), "stands included in analysis")
  ) +
  theme_light() + 
  theme(
    legend.position = "none"
    , strip.text = element_text(color = "black")
    , strip.background = element_rect(fill = "gray88")
    , axis.text.y = element_blank()
    , axis.ticks.y = element_blank()
    , panel.grid.major.y = element_blank()
    , panel.grid.minor.y = element_blank()
  )
```

### Overstory Stand Basal Area Map

```{r ba-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "basal_area_ft2_per_ac"
      , layer.name = "BA (ft2 per ac)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::cividis(n=nrow(silv_metrics)*1.1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

### Stand QMD Map

Quadratic mean diameter (QMD) is a measure of the diameter of the tree of mean basal area:

$$
\textrm{quadratic mean diameter (QMD)} = \sqrt{\frac{\sum{d_{i}^{2}}}{n}}
$$

, where $d_{i}$ is the diameter at breast height of an individual tree, and $n$ is the total number of trees. 

```{r qmd-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "qmd_in"
      , layer.name = "QMD (in)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::mako(n=nrow(silv_metrics)*1.1, direction = -1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

### Overstory DBH Distrubtion

```{r os-dbh-dist}
harvests_trees %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(dbh_cm >= ostory_min_tree_dbh_cm) %>% 
  dplyr::select(treeID, dbh_in) %>% 
  dplyr::ungroup() %>%
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = dbh_in)
  ) +
    geom_density(alpha = 0.8, fill = "navy", color = NA) + 
    labs(
      x = "DBH (in)"
      , y = "density"
      , title = paste0(
        "Overstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) SfM-derived tree DBH distribution"
      )
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

```{r os-dbh-dist2}
tpa_temp = 
  harvests_trees %>%
    sf::st_drop_geometry() %>%
    dplyr::filter(dbh_cm >= ostory_min_tree_dbh_cm) %>% 
    dplyr::select(treeID, dbh_in) %>% 
    dplyr::ungroup() %>%
    dplyr::distinct() %>% 
    dplyr::mutate(
      dbh_cl = ggplot2::cut_width(dbh_in, width = 1, center = 0.5, closed = "left")
    ) %>% 
    dplyr::group_by(dbh_cl) %>% 
    dplyr::summarise(
      n_trees = dplyr::n()
      , trees_per_ac = dplyr::n()/
        (harvests %>% sf::st_union() %>% sf::st_area() %>% as.numeric() %>% `/`(10000) %>% `*`(2.471))
      , trees_per_ha = dplyr::n()/
        (harvests %>% sf::st_union() %>% sf::st_area() %>% as.numeric() %>% `/`(10000))
    ) %>%
    dplyr::ungroup() %>% 
    dplyr::mutate(
      tot_trees_per_ac = sum(trees_per_ac)
      , tot_trees_per_ha = sum(trees_per_ha)
      , pct = trees_per_ac/tot_trees_per_ac
      , dbh_cl_lab = paste0(
        stringr::word(dbh_cl, 1, sep = fixed(",")) %>% readr::parse_number()
        , " to "
        , stringr::word(dbh_cl, -1, sep = fixed(",")) %>% readr::parse_number() %>% `-`(0.1)
      ) %>% 
      factor() %>% 
      forcats::fct_reorder(stringr::word(dbh_cl, 1, sep = fixed(",")) %>% readr::parse_number())
    )
# plot
tpa_temp %>% 
  ggplot(
    mapping = aes(
      x = dbh_cl_lab, y = trees_per_ac
      , fill=trees_per_ac
      , label = paste0(
        scales::comma(trees_per_ac, accuracy = 0.1)
        , "\n"
        , scales::percent(pct, accuracy = 0.1)
      )
    )
  ) +
    geom_col(width = 0.7) +
    geom_text(color = "black", size = 3, vjust = -0.2) +
    scale_fill_viridis_c(option = "mako", direction = -1) +
    scale_y_continuous(labels = scales::comma_format(accuracy = 1),expand = expansion(mult = c(0, .13))) +
    labs(
      x = "DBH (in) Class"
      , y = "TPA"
      , title = paste0(
        "Overstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) Trees Per Acre by Diameter Class"
      )
      , subtitle = paste0(
          "Total TPA: "
          , scales::comma(tpa_temp$tot_trees_per_ac[1], accuracy = 1)
        )
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.text.x = element_text(angle = 90, size = 10, face = "bold", vjust = 0.5, hjust = 1)
    )
```

## Height Distribution

```{r os-ht-dist}
harvests_trees %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(dbh_cm >= ostory_min_tree_dbh_cm) %>% 
  dplyr::select(treeID, tree_height_ft) %>% 
  dplyr::ungroup() %>%
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = tree_height_ft)
  ) +
    geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
    labs(
      x = "Height (ft)"
      , y = "density"
      , title = paste0(
        "Overstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) SfM-derived tree height distribution"
      )
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Understory Summary

```{r us-tpa}
# total
tpa_temp = harvests_regen_plt_diam %>% 
  sf::st_drop_geometry() %>%
  dplyr::mutate(
    dplyr::across(
      tidyselect::starts_with("diam_inch_class")
      , .fn = as.numeric
    )
  ) %>% 
  dplyr::filter(
    diam_inch_class_lower < ustory_max_tree_dbh_cm
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::distinct(plot_dir_id, regen_size_class_diam, number_of_trees_in_size_class, plot_area_ac, plot_area_ha) %>% 
  dplyr::group_by(regen_size_class_diam) %>% 
  dplyr::summarise(
    number_of_plots = dplyr::n_distinct(plot_dir_id)
    , sum_num_trees = sum(number_of_trees_in_size_class)
    , sampled_area_ac = sum(plot_area_ac)
    , sampled_area_ha = sum(plot_area_ha)
  ) %>% 
  dplyr::mutate(
    size_class_trees_per_ac = sum_num_trees/sampled_area_ac
    , size_class_trees_per_ha = sum_num_trees/sampled_area_ha
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    tot_trees_per_ac = sum(size_class_trees_per_ac)
    , tot_trees_per_ha = sum(size_class_trees_per_ha)
    , pct = size_class_trees_per_ac/tot_trees_per_ac
    , dbh_cl_lab = paste0(
        stringr::word(regen_size_class_diam, 1, sep = fixed("-")) %>% readr::parse_number()
        , " to "
        , stringr::word(regen_size_class_diam, -1, sep = fixed("-")) %>% readr::parse_number() %>% `-`(0.1)
      ) %>% 
      factor() %>% 
      forcats::fct_reorder(stringr::word(regen_size_class_diam, 1, sep = fixed("-")) %>% readr::parse_number())
  )
# plot
tpa_temp %>% 
  ggplot(
    mapping = aes(
      x = dbh_cl_lab, y = size_class_trees_per_ac
      , fill=size_class_trees_per_ac
      , label = paste0(
        scales::comma(size_class_trees_per_ac, accuracy = 0.1)
        , "\n"
        , scales::percent(pct, accuracy = 0.1)
      )
    )
  ) +
    geom_col(width = 0.7) +
    geom_text(color = "black", size = 3, vjust = -0.2) +
    scale_fill_viridis_c(option = "mako", direction = -1) +
    scale_y_continuous(labels = scales::comma_format(accuracy = 1),expand = expansion(mult = c(0, .13))) +
    labs(
      x = "DBH (in) Class"
      , y = "TPA"
      , title = paste0(
        "Understory (< "
        , scales::comma(ustory_max_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) Trees Per Acre by Diameter Class"
      )
      , subtitle = paste0(
          "Total TPA: "
          , scales::comma(tpa_temp$tot_trees_per_ac[1], accuracy = 1)
        )
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.text.x = element_text(angle = 90, size = 10, face = "bold", vjust = 0.5, hjust = 1)
    )
  
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Define Functions for Stand Summary

Create plots to combine for report

1) orthomosaic + chm + tree tops + stand plot

```{r ortho-plt-fn}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
ortho_plt_fn = function(my_suid){

# convert to stars
  ortho_st = ortho_rast %>%  
    terra::subset(subset = c(1,2,3)) %>%
    terra::crop(
      # stand %>% 
      harvests %>% dplyr::filter(suid==my_suid) %>% 
        sf::st_buffer(20) %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        terra::vect()
    ) %>% 
    terra::aggregate(fact = 4, fun = "mean", na.rm = T) %>% 
    stars::st_as_stars()
  
  # convert to rgb
  ortho_rgb <- stars::st_rgb(
    ortho_st[,,,1:3]
    , dimension = 3
    , use_alpha = FALSE
    # , stretch = "histogram"
    , probs = c(0.005, 0.995)
    , stretch = "percent"
  )
  # ggplot
  plt_rgb <- ggplot() +
    stars::geom_stars(data = ortho_rgb[]) +
    scale_fill_identity(na.value = "transparent") + # !!! don't take this out or RGB plot will kill your computer
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(
      x = ""
      , y = ""
    ) +
    theme_void()
  
  # chm
  chm = chm_rast %>% 
    terra::crop(
      harvests %>% dplyr::filter(suid==my_suid) %>% 
      terra::vect()
    ) %>% 
    terra::mask(
      harvests %>% dplyr::filter(suid==my_suid) %>% 
      terra::vect()
    ) %>% 
    terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
    `*`(3.28) %>% # transform to feet
    as.data.frame(xy=T) %>% 
    rename(f=3)
  
  # return(plt_rgb)
  # combine all plot elements
  plt_combine = plt_rgb +
    # # understory
    # geom_tile(
    #   data = chm %>% 
    #     dplyr::filter(
    #       f <
    #       pred_mod_best %>%
    #         dplyr::filter(estimate_in < ustory_max_tree_dbh_cm/2.54) %>% 
    #         dplyr::pull(tree_height_ft) %>% max()
    #     )
    #   , mapping = aes(x = x, y = y)
    #   , fill = "gray"
    #   , alpha = 0.4
    #   , na.rm = T
    # ) +
    # overstory
    ggnewscale::new_scale_fill() +
    geom_tile(
      data = chm %>% 
        dplyr::filter(
          f >=
          pred_mod_best %>%
            dplyr::filter(estimate_in >= ostory_min_tree_dbh_cm/2.54) %>% 
            dplyr::pull(tree_height_ft) %>% min()
        )
      , mapping = aes(x = x, y = y, fill = f)
      , na.rm = T
    ) +
    scale_fill_viridis_c(option="plasma", alpha = 0.6, breaks = scales::extended_breaks(n=6), na.value = "transparent") +
    geom_sf(
      data = harvests %>% dplyr::filter(suid==my_suid)
      , alpha = 0
      , lwd = 1.5
      , color = "#b22222"
    ) +
    # geom_sf(
    #   data = treetops_sf_with_dbh %>% 
    #     sf::st_intersection(harvests %>% dplyr::filter(suid==my_suid))
    #   , color = "#330099"
    #   , shape = "."
    # ) +
    labs(
      fill = "overstory\ncanopy ht. (ft)"
      # , title = "Trees identified within stand boundary"
      # , title = "<span><span style='color:#330099;'><b><i>Trees</i></b></span> identified within <span style='color:#b22222;'><b><i>stand</i></b></span> boundary</span>"
      , title = paste0(
        "Overstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH) within stand boundary"
      )
      # , subtitle = paste0(
      #   "trees under "
      #   , scales::comma((ustory_max_tree_dbh_cm/2.54),accuracy = 0.1)
      #   , " in. DBH shown in gray"
      # )
    ) +
    theme(
      legend.position = "top" # c(0.5,1)
      , legend.direction = "horizontal"
      , legend.margin = margin(0,0,-1,0)
      , legend.text = element_text(size = 8)
      , legend.title = element_text(size = 8)
      # , plot.title = ggtext::element_markdown(size = 10, hjust = 0.5)
      , plot.title = element_text(size = 10, hjust = 0.5, face = "bold")
      , plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic")
    )
  return(plt_combine)
}
# ortho_plt_fn(harvests$suid[3])
```

```{r, include=FALSE, eval=F}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
  xxxxxortho_plt_fn = function(stand,ortho,points,chm){
      # plot ortho
      ortho %>% 
        terra::crop(
          stand %>% 
            sf::st_buffer(20) %>% 
            sf::st_bbox() %>% 
            sf::st_as_sfc() %>% 
            terra::vect()
        ) %>% 
        terra::aggregate(fact = 4, fun = "mean", na.rm = T) %>% 
        terra::plotRGB(
          r = 1, g = 2, b = 3, stretch = "lin", colNA = "transparent"
          , mar = c(4,0,2,0)
          , main = "Trees (blue) identified within stand boundary (red)"
          , cex.main = 0.8
        )
      # plot chm
      chm %>% 
        `*`(3.28) %>% # transform to feet
        terra::crop(stand %>% terra::vect()) %>% 
        terra::mask(stand %>% terra::vect()) %>%
        terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
        terra::plot(
          col = viridis::plasma(50), colNA = NA, alpha = 0.6, add = T
          , plg = list(
            title = "canopy ht. (ft)"
            , title.cex = 0.8
            , title.adj = 0
          )
        )
      # legend(legend = 2,title = "canopy ht. (ft)")
      
      # plot stand
        terra::plot(
          stand %>% terra::vect()
          , col = NA, border = "firebrick", lwd = 2.5
          , add = T
        )
      # plot points
      terra::plot(
        terra::crop(points %>% terra::vect(), stand %>% terra::vect())
        , pch = "."
        , col = "blue"
        , add = T
      )
      
  }
# ortho_1 = function(){ortho_plt_fn(
#     stand = harvests[1,]
#     , ortho = ortho_rast
#     , points = treetops_sf_with_dbh
#     , chm = chm_rast
#   )}
# ortho_1()
```

2) dbh distribution plot

```{r dbh-dist-fn}
dbh_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(
      suid == my_suid
      & dbh_cm >= ostory_min_tree_dbh_cm
    ) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, dbh_in) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = dbh_in)
    ) +
      geom_density(alpha = 0.8, fill = "navy", color = NA) + 
      labs(
        x = "DBH (in)"
        , y = "density"
        , title = paste0(
          "Overstory (>= "
          , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
          , " in. DBH) DBH distribution"
        )
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# dbh_dist_fn(harvests$suid[1])
```

3) height distribution plot

```{r ht-dist-fn}
ht_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(
      suid == my_suid
      & dbh_cm >= ostory_min_tree_dbh_cm
    ) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, tree_height_ft) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = tree_height_ft)
    ) +
      geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
      labs(
        x = "Height (ft)"
        , y = "density"
        , title = paste0(
          "Overstory (>= "
          , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
          , " in. DBH) height distribution"
        )
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# ht_dist_fn(harvests$suid[1])
```

4) silvicultural metrics imperial table

```{r tbl-fn}
tab_imp_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ac"
      , "n_trees"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2_per_ac"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ac"
            , "n_trees"
            , "trees_per_ac"
            , "mean_tree_height_ft"
            , "loreys_height_ft"
            , "mean_dbh_in"
            , "qmd_in"
            , "basal_area_ft2_per_ac"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ac)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ac^{-1}$")
            , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
            , latex2exp::TeX("Mean DBH (in)", output = "character")
            , latex2exp::TeX("QMD (in)", output = "character")
            , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = paste0(
        "Silvicultural Metrics (imperial)"
        , "\nOverstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm/2.54,accuracy = 0.1)
        , " in. DBH)"
        )
      , face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}

# tab_imp_fn(harvests$suid[2])

```

5) silvicultural metrics metric table

```{r tbl-fn2}
tab_met_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ha"
      , "n_trees"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2_per_ha"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ha"
            , "n_trees"
            , "trees_per_ha"
            , "mean_tree_height_m"
            , "loreys_height_m"
            , "mean_dbh_cm"
            , "qmd_cm"
            , "basal_area_m2_per_ha"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ha)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ha^{-1}$")
            , latex2exp::TeX("Mean Tree Height (m)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (m)", output = "character")
            , latex2exp::TeX("Mean DBH (cm)", output = "character")
            , latex2exp::TeX("QMD (cm)", output = "character")
            , latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = paste0(
        "Silvicultural Metrics (metric)"
        , "\nOverstory (>= "
        , scales::comma(ostory_min_tree_dbh_cm,accuracy = 0.1)
        , " cm DBH)"
        )
      , face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}

# tab_met_fn(harvests$suid[2])
```

6) inset map plot

```{r inset-map}
plt_inset_map_fn <- function(my_suid){
  (
    ggplot() +
      geom_sf(data = bhef_boundary, alpha = 0, lwd = 1, color = "black") +
      geom_sf(data = harvests, alpha = 0, lwd = 0.4, color = "gray") +
      geom_sf(
        data = harvests %>% dplyr::filter(suid == my_suid)
        , fill = "firebrick"
        , alpha = 0.3
        , lwd = 1
        , color = "firebrick"
      ) +
      # geom_sf_label(
      #   data = rx_fire
      #   , aes(label = unit)
      #   , label.size = NA
      #   , alpha = 0
      #   , size = 2.5
      # ) +
      xlab("") +
      ylab("") +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_text(size = 3.5)
        , axis.text.x = element_text(angle = 35)
        , panel.border = element_blank()
      )

  )
}
# plt_inset_map_fn(harvests$suid[4])
```

7) contour + elevation map plot

```{r elev-plt-inset}
# sequence by 5's to get 5m breaks list for contour
contour_int_m <- 5
seq_brk <- seq(0, 40000, contour_int_m)
# create contour data
elev_contour <- dtm_rast %>% 
  stars::st_as_stars() %>% 
  stars::st_contour(
    na.rm = TRUE
    , contour_lines = FALSE
    , breaks = # classInt::classIntervals(na.omit(as.vector(x[[1]])))$brks
        seq_brk[
          seq_brk > min(na.omit(as.vector(dtm_rast[[1]]))) - contour_int_m
          & seq_brk < max(na.omit(as.vector(dtm_rast[[1]]))) + contour_int_m
        ]
  )

plt_contour_map_fn <- function(my_suid){
  (
    #########################################
    #########################################
    # map of unit w/ contour
    #########################################
    #########################################
    # plot
    ggplot() +
      geom_tile(
        data = dtm_rast %>% 
          terra::crop(
            harvests %>% dplyr::filter(suid == my_suid) %>% 
              sf::st_buffer(dist = 20) %>% 
              sf::st_bbox() %>% 
              sf::st_as_sfc() %>% 
              terra::vect()
          ) %>% 
          terra::aggregate(fact=2) %>% 
          `*`(3.28) %>% # transform to feet
          as.data.frame(xy=T) %>% 
          dplyr::rename(elev = 3)
        , mapping = aes(x=x,y=y,fill=elev)
      ) +
      geom_sf(data = elev_contour %>% 
        sf::st_crop(
          harvests %>% dplyr::filter(suid == my_suid) %>% 
          sf::st_buffer(dist = 20) %>% 
          sf::st_bbox()
        )
        , alpha = 0
        , lwd = 0.5
        , color = "gray30"
        , show.legend = FALSE
      ) +
      geom_sf(data = harvests %>% dplyr::filter(suid == my_suid)
          , alpha = 0
          , lwd = 1.5
          , color = "firebrick"
          , show.legend = FALSE
      ) +
      scale_fill_viridis_c(
        option = "viridis"
        , alpha = 0.9
        , na.value = "transparent"
        , breaks = scales::extended_breaks(n = 6)
        , labels = scales::comma
      ) +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      labs(
        subtitle = paste0(
          as.character(round(contour_int_m * 3.28, 1))
          , " ft ("
          , as.character(contour_int_m)
          , " m) contours"
        )
        , fill = "elev. (ft)"
        , x = ""
        , y = ""
      ) +
      theme_void() +
      theme(
        legend.position = "top"
        , legend.direction = "horizontal"
        , legend.margin = margin(0,0,-1,0)
        # , legend.box.margin = margin(-1,0,2,0)
        , legend.text = element_text(size = 4, angle = 25, hjust = 0.7)
        , legend.title = element_text(size = 4)
        , plot.subtitle = element_text(
            size=5, face="italic", color="gray20", margin = margin(0,0,0,0)
            , hjust = 0.5
        )
      )
  )
}
# plt_contour_map_fn(harvests$suid[1])
```

8) regen by size class

```{r stand-regen-class}
plt_regen_dist_fn <- function(my_suid){
  dta = harvests_regen_plt_diam %>% 
    sf::st_drop_geometry() %>%
    dplyr::mutate(
      dplyr::across(
        tidyselect::starts_with("diam_inch_class")
        , .fn = as.numeric
      )
    ) %>% 
    dplyr::filter(
      suid == my_suid
      & diam_inch_class_lower < ustory_max_tree_dbh_cm
    ) %>% 
    dplyr::group_by(suid,regen_size_class_diam) %>% 
    dplyr::summarise(
      number_of_plots = dplyr::n()
      , sum_num_trees = sum(number_of_trees_in_size_class)
      , sampled_area_ac = sum(plot_area_ac)
    ) %>% 
    dplyr::group_by(suid) %>% 
    dplyr::mutate(
      trees_per_ac = sum_num_trees/sampled_area_ac
      , stand_trees_per_ac = sum(sum_num_trees)/max(sampled_area_ac)
      , regen_size_class_diam = regen_size_class_diam %>% factor() %>% forcats::fct_rev()
    ) %>% 
    dplyr::ungroup() %>% 
    dplyr::mutate(
      stand_lab = paste0(
          # suid
          "\nTotal TPA: "
          , scales::comma(stand_trees_per_ac, accuracy = 1)
      ) 
      # factor() %>% forcats::fct_reorder(desc(stand_trees_per_ac))
    )
    ## plt
    ggplot(
      data = dta
      , mapping = aes(
        x = trees_per_ac, y = regen_size_class_diam
        , fill=trees_per_ac
        , label = scales::comma(trees_per_ac, accuracy = 0.1)
      )
    ) +
      geom_col(
        width = 0.7
      ) +
      geom_text(
        color = "black", size = 2
        , hjust = -0.1
      ) +
      scale_fill_viridis_c(option = "mako", direction = -1) +
      scale_x_continuous(expand = expansion(mult = c(0, .13))) +
      # facet_wrap(facets = vars(stand_lab), ncol = 3) +
      labs(
        fill = ""
        , y = "DBH (in) Class"
        , x = "Trees per Acre"
        , title = paste0(
          "Understory (< "
          , scales::comma(ustory_max_tree_dbh_cm/2.54,accuracy = 0.1)
          , " in. DBH) Trees Per Acre"
        )
        # , subtitle = "by harvest unit (with FACTS ID)"
        , subtitle = dta$stand_lab[1]
      ) +
      theme_light() +
      theme(
        legend.position = "none"
        , axis.text.x = element_blank()
        , axis.text.y = element_text(color = "black",size=12, face = "bold")
        , axis.ticks.x = element_blank()
        # , strip.text = element_text(color = "black", size = 12)
        # , strip.background = element_rect(fill = "gray88")
      )
}
# plt_regen_dist_fn(harvests$suid[2])
```

81) build report

```{r report-fn}
report_fn = function(my_suid){
  # generate plots
  ortho_plt_temp = ortho_plt_fn(my_suid) # ortho_plt_fn(harvests$suid[4])
  dbh_dist_temp = dbh_dist_fn(my_suid) + # dbh_dist_fn(harvests$suid[4])
    theme(
      plot.background = element_rect(colour = "gray", fill=NA, size=1)
      # , plot.title = element_text(size = 9)
      # , plot.subtitle = element_text(size = 7)
    )
  ht_dist_temp = ht_dist_fn(my_suid) + # ht_dist_fn(harvests$suid[4]) 
    theme(
      plot.background = element_rect(colour = "gray", fill=NA, size=1)
      # , plot.title = element_text(size = 9)
      # , plot.subtitle = element_text(size = 7)
    )
  plt_inset_map_temp = plt_inset_map_fn(my_suid) # plt_inset_map_fn(harvests$suid[4])
  plt_contour_map_temp = plt_contour_map_fn(my_suid) # plt_contour_map_fn(harvests$suid[4])
  tab_imp_temp = tab_imp_fn(my_suid) # tab_imp_fn(harvests$suid[4])
  tab_met_temp = tab_met_fn(my_suid) # tab_met_fn(harvests$suid[4]) 
  plt_regen_dist_temp = plt_regen_dist_fn(my_suid) +
    theme(plot.background = element_rect(colour = "gray", fill=NA, size=1))
  # plot layout
  # area(t,l,b,r)
  layout_temp = c(
    # map
    area(1, 2, 2, 2)
    # map inset
    , area(1, 1, 1, 1)
    # contour
    , area(2, 1, 2, 1)
    # # tables
    , area(4, 1, 5, 1)
    , area(6, 1, 7, 1)
    # space
    , area(3, 1, 3, 2)
    # # # distributions 
    , area(5, 2, 5, 2)
    , area(6, 2, 6, 2)
    , area(7, 2, 7, 2)
  )
  # check the layout
  # plot(layout_temp)
  
  # patchwork 
  rpt = 
    ortho_plt_temp +
    plt_inset_map_temp + 
    plt_contour_map_temp + 
    tab_imp_temp + 
    tab_met_temp + 
    patchwork::plot_spacer() +
    dbh_dist_temp + ht_dist_temp + plt_regen_dist_temp +
    patchwork::plot_layout(
      design = layout_temp
      , heights = c(0.7, 0.8, 0.3, 1,1,1,1)
      , widths = c(0.5,1)
    )
  # export
  ggplot2::ggsave(
    filename = paste0(delivery_dir, "/suid_", my_suid, ".pdf")
    , plot = rpt
    , device = "pdf"
    , width = 8.5, height = 11, units = "in"
  )
  return(rpt)
}
```

## Harvest Unit Summary Reports

```{r print-reports, results='hide', fig.show='asis', fig.height=11}
# generate and print reports
harvests$suid %>% 
  # .[1] %>% 
  purrr::map(report_fn) %>% 
  print()
```

```{r pdf-comb, results='hide'}
# combine all pdfs together
# combine with pdftools
pdftools::pdf_combine(
  sort(list.files(delivery_dir, pattern = "\\.pdf$", full.names = TRUE))
  , output = paste0(
    delivery_dir
    , "/BHEF_UAS_202306_report_"
    , gsub("-", "", Sys.Date())
    , ".pdf"
  )
)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
# check_ls_size_fn(ls()) %>% dplyr::slice_head(n=20)
remove(
  elev_contour
  , seq_brk
  , ortho_plt_fn
  , report_fn
  , plt_contour_map_fn
  , tab_imp_fn
  , tab_met_fn
  , plt_inset_map_fn
  , ht_dist_fn
)
remove(list = ls()[grep("_temp",ls())])
gc()
```


```{r include=FALSE, eval=FALSE}
library(patchwork)
# area(t,l,b,r)
layout_temp = c(
    # map
    area(1, 2, 2, 2)
    # map inset
    , area(1, 1, 1, 1)
    # contour
    , area(2, 1, 2, 1)
    # # tables
    , area(4, 1, 5, 1)
    , area(6, 1, 7, 1)
    # space
    , area(3, 1, 3, 2)
    # # # distributions 
    , area(5, 2, 5, 2)
    , area(6, 2, 6, 2)
    , area(7, 2, 7, 2)
  )
# check the layout
  plot(layout_temp)
  
  
```

