# Regeneration Field Validation Data

The 2023-06 BHEF field surveys used $\frac{1}{400}$ acre (10.117 m^2^) plots with a 5.89 ft (1.795 m) radius for regeneration sampling. To compare the 

The F-score incorporates true positive, commission, and omission rates to determine how well the UAS detected trees represent the field-based stem mapped trees. As a measure of predictive performance, the highest possible value of an F-score is 1.0, indicating perfect precision and recall, and the lowest possible value is 0, if either precision or recall are zero.

$$
\textrm{F-score} = 2 \times \frac{\bigl(\frac{TP}{TP+Om} \times \frac{TP}{TP+Co} \bigr)}{\bigl(\frac{TP}{TP+Om} + \frac{TP}{TP+Co} \bigr)}
$$

## Load Field Data

Load the regeneration data, set parameters for processing, and map

```{r regen-dta-ld, results='hide'}
# load Field Validation Data
field_regen = readr::read_csv(
    file = "../data/field_data/Voodoo_regen_data.csv"
  ) %>% 
  dplyr::rename_with(~ .x %>%
      # replace all non alphanumeric with _
      stringr::str_replace_all("[^[:alnum:] ]+", "_") %>%
      # remove any _ at the end of the string
      stringr::str_remove("[_]$") %>% 
      tolower()
  ) %>% 
  dplyr::filter(
    !is.na(plot_id)
    & !is.na(utm_x)
    & !is.na(utm_y)
    & !is.na(regen_size_class)
    & regen_size_class != ""
    & !is.na(number_of_trees_in_size_class)
    & !is.na(average_tree_height_feet)
    & number_of_trees_in_size_class > 0
    & average_tree_height_feet > 0
  ) %>% 
  dplyr::mutate(
    regen_size_class = regen_size_class %>% stringr::str_squish() %>% tolower()
  ) %>% 
  # remove duplicates
  dplyr::group_by(plot_id, regen_size_class, select_the_regen_plot_center_north_or_south) %>% 
  dplyr::filter(dplyr::row_number()==1) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    plot_dir_id = paste0(
      stringr::str_pad(plot_id, width = 4, side = "left", pad = "0")
      , "-"
      , tolower(select_the_regen_plot_center_north_or_south)
    )
  ) %>% 
  dplyr::relocate(plot_dir_id) %>% 
  sf::st_as_sf(
    coords = c("utm_x", "utm_y")
    , crs = sf::st_crs(treetops_sf_with_dbh)
    , remove=F
  ) %>% 
  # keep only points that are within the uas flight boundary
  sf::st_intersection(las_ctg_dta)

  # field_regen %>%
  #   sf::st_drop_geometry() %>%
  #   dplyr::summarize(
  #     n = dplyr::n()
  #     , unique_plot = dplyr::n_distinct(plot_dir_id)
  #     , unique_stuff = dplyr::n_distinct(plot_id, regen_size_class, select_the_regen_plot_center_north_or_south)
  #   )
  
  
```

```{r regen-dta-str}
field_regen %>% 
  dplyr::glimpse()
```

Regeneration field data map

```{r regen-dta-map, results='asis', fig.height = 5}
field_regen %>% 
  sf::st_drop_geometry() %>% 
  dplyr::distinct(plot_dir_id, utm_x, utm_y) %>% 
  sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
  dplyr::mutate(x=1) %>% 
mapview::mapview(col.regions = "tan", layer.name = "regen plot", alpha.regions = 0.6) +
mapview::mapview(
  harvests
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
)
```

## Build UAS-Field Validation Data

### True Positive Identification

The UAS detected and stem-mapped tree pairs identified in this filtering process (detailed above) were considered true positive ($TP$) detections.

```{r regen-dta-tp, results='hide'}
## BUFFER THE UAS TREES AND SPATIALLY MATCH FIELD TREES BASED ON THAT BUFFER   
## get FIELD trees within radius OF UAS TREES
  potential_tree_pairs_temp = treetops_sf_with_dbh %>% 
    dplyr::select(treeID, tree_height_m) %>% 
    # buffer point
    sf::st_buffer(max_dist_m) %>% 
    # spatial join with all FIELD tree points
    sf::st_join(
      field_overstory %>% 
        dplyr::filter(
          field_tree_height_m >= min_tree_height_m
          & field_dbh_cm >= min_tree_dbh_cm
        ) %>% 
        dplyr::select(
          field_plottree_id, field_tree_height_m
          , tree_utm_x, tree_utm_y
        )
      , join = st_intersects
      , left = F # performs inner join to only keep uas trees with a match
    ) %>% 
    # calculate height difference
    dplyr::mutate(
      height_diff_pct = abs(tree_height_m-field_tree_height_m)/field_tree_height_m
    ) %>% 
    # removes tree pairs that are outside of the allowable error
    dplyr::filter(height_diff_pct <= max_height_error_pct) %>% 
    dplyr::relocate(treeID, field_plottree_id)

## apply pair selection criteria if there are potential tree pairs
  if(nrow(potential_tree_pairs_temp)>0){
    ## calculate row by row distances and height differences
      potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
        # this is the position of the uas tree
        sf::st_centroid() %>% 
        sf::st_set_geometry("geom1") %>% 
        dplyr::bind_cols(
          potential_tree_pairs_temp %>% 
            sf::st_drop_geometry() %>% 
            dplyr::select("tree_utm_x", "tree_utm_y") %>% 
            # this is the position of the field tree
            sf::st_as_sf(
              coords = c("tree_utm_x", "tree_utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)
            ) %>% 
            sf::st_set_geometry("geom2")
        ) %>% 
        dplyr::mutate(
          distance_m = sf::st_distance(geom1, geom2, by_element = T) %>% as.numeric()
        ) %>% 
        sf::st_drop_geometry() %>% 
        dplyr::select(-c(tree_utm_x, tree_utm_y, geom2))
    
    ## define function to select the best tree pair
      select_best_tree_pair_fn <- function(df) {
       df %>% 
        dplyr::group_by(field_plottree_id) %>% 
        dplyr::arrange(field_plottree_id, height_diff_pct, distance_m, desc(tree_height_m), treeID) %>% 
        dplyr::mutate(
          # at the field tree level...the number of field trees
          n_uas_trees = dplyr::n()
          # at the field tree level...
            # the closest uas tree in height tie breaker distance, uas_tree_height_m, id
          , rank_within_field_tree = dplyr::row_number()
        ) %>% 
        dplyr::group_by(treeID) %>% 
        dplyr::arrange(treeID, height_diff_pct, distance_m, desc(tree_height_m), field_plottree_id) %>% 
        dplyr::mutate(
          # at the uas tree level...the number of uas trees
          n_field_trees = dplyr::n()
          # at the field tree level...
            # the closest uas tree in height tie breaker distance, uas_tree_height_m, id
          , rank_within_uas_tree = dplyr::row_number()
        ) %>% 
        dplyr::ungroup() %>% 
        # select the uas-field tree pair with the minimum height difference
        dplyr::filter(
          rank_within_field_tree == 1
          & rank_within_uas_tree == 1
        ) %>% 
        # remove columns
        dplyr::select(
          -c(tidyselect::starts_with("rank_"), tidyselect::starts_with("n_"))
        )
      }
    
    ## first filter for tree pairs
      true_positive_trees = select_best_tree_pair_fn(potential_tree_pairs_temp)
    
    ##remove matches from potential tree pairs 
      potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
        dplyr::filter(
          !(treeID %in% true_positive_trees$treeID)
          & !(field_plottree_id %in% true_positive_trees$field_plottree_id)
        )
      
    ## keep filtering for best pair until no unique pairs remain
      while(nrow(potential_tree_pairs_temp)>0) {
        # keep filtering for best pair until no unique pairs remain
        true_positive_trees = true_positive_trees %>% 
          dplyr::bind_rows(
            select_best_tree_pair_fn(potential_tree_pairs_temp)
          )
        #remove matches from potential tree pairs 
        potential_tree_pairs_temp = potential_tree_pairs_temp %>% 
          dplyr::filter(
            !(treeID %in% true_positive_trees$treeID)
            & !(field_plottree_id %in% true_positive_trees$field_plottree_id)
          )
      }
    ## rename columns and flag
      true_positive_trees = true_positive_trees %>% 
        dplyr::rename(
          uas_tree_height_m = tree_height_m
          , uas_tree_id = treeID
          , field_uas_distance_m = distance_m
        ) %>% 
        dplyr::mutate(
          field_uas_group = "true positive"
        )
  }else{ # if there are spatially matched trees
    true_positive_trees = dplyr::tibble(
      uas_tree_id = as.character(NA)
      , field_plottree_id = as.character(NA)
      , uas_tree_height_m = as.numeric(NA)
      , field_tree_height_m = as.numeric(NA)
      , height_diff_pct = as.numeric(NA)
      , field_uas_distance_m = as.numeric(NA)
    )
  }
  remove(potential_tree_pairs_temp)
  gc()
```

True positive detections

```{r regen-dta-tp-data}
#check data
  true_positive_trees %>%
    dplyr::summarise(
      n_rows = n()
      , unique_uas_trees = n_distinct(uas_tree_id)
      , unique_field_trees = n_distinct(field_plottree_id)
      , unique_pairs = n_distinct(field_plottree_id, uas_tree_id)
    ) %>% 
    kableExtra::kbl(caption = "results of true positive stem-uas tree matching") %>% 
    kableExtra::kable_styling()
```

### Combine with Commission and Omission

To determine UAS detected tree commissions (i.e. UAS detected trees within the overstory plot for which there was no stem-mapped tree pair; $Co$) this analysis used the 2023-06 BHEF overstory field survey plot center and plot radius of 11.35 m. UAS detected trees within this radius with an [estimated DBH](#local_mod) over 5 in (12.69 cm) that did not have a matched stem-mapped tree pair were considered commissions ($Co$).

Omissions ($Om$) are stem-mapped trees without a UAS detected tree match.

```{r regen-dta-comb, results='hide'}
field_uas_comparison = dplyr::bind_rows(
  ## true positive
    true_positive_trees
  ## omission
    , field_overstory %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(
        plot_id, field_plottree_id, field_tree_height_m
      ) %>% 
      dplyr::anti_join(
        true_positive_trees
        , by = dplyr::join_by(field_plottree_id)
      ) %>% 
      dplyr::mutate(
        field_uas_group = "omission"
      )
  ## commission
    , field_overstory %>% 
      sf::st_drop_geometry() %>% 
      # get field plot center locations
      dplyr::distinct(plot_id, utm_x, utm_y) %>% 
      sf::st_as_sf(coords = c("utm_x", "utm_y"), crs = sf::st_crs(treetops_sf_with_dbh)) %>% 
      # buffer plot center with plot area 11.35m radius + mean error in true positive trees
      # sf::st_buffer(11.35+mean(true_positive_trees$field_uas_distance_m, na.rm = T)) %>% 
      # buffer plot center with plot area 11.35m radius
      sf::st_buffer(11.35) %>% 
      # join with uas tree points
      sf::st_join(
        treetops_sf_with_dbh %>% 
          dplyr::filter(
            !treeID %in% true_positive_trees$uas_tree_id
            & dbh_cm >= min_tree_dbh_cm
            & tree_height_m >= min_tree_height_m
          ) %>% 
          dplyr::select(treeID) %>% 
          dplyr::rename(uas_tree_id=treeID)
        , join = st_intersects
      ) %>% 
      sf::st_drop_geometry() %>% 
      dplyr::mutate(
        field_uas_group = "commission"
      )
) %>% 
  # attach uas data
  dplyr::left_join(
    treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(treeID, tree_height_m, dbh_cm) %>% 
      dplyr::rename(
        uas_tree_id = treeID
        , uas_tree_height_m = tree_height_m
        , uas_dbh_cm = dbh_cm
      )
    , by = dplyr::join_by(uas_tree_id)
  ) %>% 
  # attach field data
  dplyr::left_join(
    field_overstory %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(field_plottree_id, field_tree_height_m, field_dbh_cm, plot_id)
    , by = dplyr::join_by(field_plottree_id)
  ) %>% 
  # update data
  dplyr::mutate(
    uas_tree_height_m = uas_tree_height_m.y
    , field_tree_height_m = field_tree_height_m.y
    , plot_id = dplyr::coalesce(plot_id.x, plot_id.y)
    , field_uas_group = factor(
      field_uas_group
      , ordered = T
      , levels = c(
        "true positive"
        , "commission"
        , "omission"
      )
    ) %>% forcats::fct_rev()
    , dbh_diff_cm = uas_dbh_cm - field_dbh_cm
    , tree_height_diff_m = uas_tree_height_m - field_tree_height_m
    , dbh_diff_pct = dbh_diff_cm/field_dbh_cm
    , height_diff_pct = tree_height_diff_m/field_tree_height_m
    , abs_dbh_diff_pct = abs(dbh_diff_pct)
    , abs_height_diff_pct = abs(height_diff_pct)
    , plot_lab = stringr::str_pad(plot_id, width = 4, side = "left", pad = "0")
  ) %>% 
  dplyr::relocate(field_uas_group) %>% 
  dplyr::select(-c(tidyselect::ends_with(".x"), tidyselect::ends_with(".y"))) %>% 
  # convert to imperial units
  calc_imperial_units_fn()

## write
write.csv(field_uas_comparison, paste0(delivery_dir,"/field_uas_comparison_data.csv"), row.names = F)

## aggregate data to plot level
field_uas_comparison_plot =
  field_uas_comparison %>% 
  dplyr::count(plot_id, plot_lab, field_uas_group) %>% 
  dplyr::group_by(plot_id, plot_lab) %>% 
  dplyr::mutate(
    pct = n/sum(n)
    , tot = sum(n)
    , tp_pct=max(ifelse(field_uas_group=="true positive",pct,0)) 
  ) %>% 
  dplyr::ungroup() %>% 
  # attach f score
  dplyr::inner_join(
    field_uas_comparison %>% 
      dplyr::count(plot_id, field_uas_group) %>% 
      dplyr::mutate(field_uas_group = dplyr::case_when(
        field_uas_group == "true positive" ~ "tp"
        , field_uas_group == "commission" ~ "co"
        , field_uas_group == "omission" ~ "om"
      )) %>% 
      tidyr::pivot_wider(
        names_from = field_uas_group
        , values_from = n
        , values_fill = 0
      ) %>% 
      dplyr::mutate(
        plot_f_score = dplyr::coalesce(
          2 * ( (tp/(tp+om)) * (tp/(tp+co)) ) / ( (tp/(tp+om)) + (tp/(tp+co)) )
          , 0
        )
      ) %>% 
      dplyr::select(plot_id, plot_f_score)
    , by = dplyr::join_by(plot_id)
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Summary of Validation Data

### True Positive, Commission, Ommission

Summary of tree true positive ($TP$), commission ($Co$), and omission ($Om$) detection by BHEF overstory plot.

```{r regen-dta-sum-grp}
field_uas_comparison_plot %>% 
ggplot(
  mapping = aes(x = pct, y = reorder(plot_lab, tp_pct), fill=field_uas_group, group=field_uas_group)
) +
  geom_col(
    width = 0.7, alpha=0.8
  ) +
  geom_text(
    mapping = aes(
        label = scales::percent(ifelse(pct>=0.12,pct,NA), accuracy = 1)
        , fontface = "bold"
      )
    , position = position_stack(vjust = 0.5)
    , color = "black", size = 2.3
  ) +
  scale_fill_viridis_d(option = "cividis") +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(
    fill = ""
    , y = ""
    , x = "Percent of Trees"
    , title = "UAS and Stem-Mapped Tree Validation Summary"
    , subtitle = "by overstory plot"
  ) +
  theme_light() +
  theme(
    legend.position = "top"
    , legend.direction  = "horizontal"
    , legend.title = element_text(size=7)
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_blank()
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  ) +
  guides(
    fill = guide_legend(reverse = T, override.aes = list(alpha = 0.9))
  )  
      
```

### F-score

```{r regen-dta-sum-f}
field_uas_comparison_plot %>% 
  dplyr::distinct(plot_lab, plot_f_score) %>% 
ggplot(
  mapping = aes(x = plot_f_score, y = reorder(plot_lab, plot_f_score), fill=plot_f_score, label = scales::comma(plot_f_score, accuracy = 0.01))
) +
  geom_vline(
    xintercept = field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median()
    , color = "gray44"
    , linetype = "dashed"
  ) +
  geom_col(
    width = 0.7, alpha=0.8
  ) +
  geom_text(
    color = "black", size = 3.5
    , hjust = -0.1
  ) +
  annotate(
    "text"
    , x = field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median()
    , y = 0 
    , label = paste0(
      "median: "
      , field_uas_comparison_plot %>% dplyr::distinct(plot_lab, plot_f_score) %>% 
      dplyr::pull(plot_f_score) %>% median() %>% scales::comma(accuracy = 0.01)
    )
    , hjust = -0.1
    , vjust = -1
  ) +
  scale_fill_viridis_c(option = "mako", direction = -1) +
  labs(
    fill = ""
    , y = ""
    , x = "F-Score"
    , title = "UAS and Stem-Mapped Tree F-Score Summary"
    , subtitle = "by overstory plot"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
    , axis.title.x = element_text(size=10, face = "bold")
    , axis.title.y = element_blank()
    , axis.text.x = element_blank()
    , axis.text.y = element_text(color = "black",size=10, face = "bold")
    , axis.ticks.x = element_blank()
  )
    
```

### Height vs. DBH of $Tp$, $Co$, $Om$

```{r regen-dta-sum-htdbh}
field_uas_comparison %>% 
  dplyr::mutate(
    dbh_temp = dplyr::coalesce(field_dbh_in, uas_dbh_in)
    , ht_temp = dplyr::coalesce(field_tree_height_ft, uas_tree_height_ft)
  ) %>% 
  ggplot(
    mapping = aes(x = ht_temp, y = dbh_temp, color = field_uas_group)
  ) +
    geom_point(
      mapping = aes(shape = field_uas_group)
      , alpha=0.8
      , size=2
    ) +
    scale_color_viridis_d(option = "cividis") +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    scale_y_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      color = "detection"
      , shape = "detection"
      , y = "DBH (in)"
      , x = "Tree Ht. (ft)"
      , title = "UAS and Stem-Mapped Tree Validation Summary"
      , subtitle = "height and DBH relationship"
    ) +
    theme_light() +
    theme(
      legend.position = "top"
      , legend.direction  = "horizontal"
      , legend.title = element_blank()
    ) +
    guides(
      color = guide_legend(reverse = T, override.aes = list(alpha = 0.9, size = 5))
      , shape = guide_legend(reverse = T, override.aes = list(alpha = 0.9, size = 5))
    )  
```

### Height and DBH Distribution $Tp$, $Co$, $Om$

```{r regen-dta-sum-htdbh-dist}
field_uas_comparison %>% 
  dplyr::mutate(
    dbh = dplyr::coalesce(field_dbh_in, uas_dbh_in)
    , height = dplyr::coalesce(field_tree_height_ft, uas_tree_height_ft)
  ) %>% 
  dplyr::select(dbh, height, field_uas_group) %>% 
  tidyr::pivot_longer(cols = -c(field_uas_group), names_to = "metric", values_to = "value") %>% 
  dplyr::group_by(field_uas_group,metric) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      field_uas_group
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = value, y = plot_lab, fill = field_uas_group)) +
    geom_violin(outlier.shape = NA) +
    geom_boxplot(width = 0.1, outlier.shape = NA) +
    facet_grid(cols = vars(metric), scales = "free_x") +
    scale_fill_viridis_d(option = "cividis") +
    scale_x_continuous(breaks = scales::extended_breaks(n=8)) +
    labs(
      fill = ""
      , y = ""
      , x = ""
      , title = "UAS and Stem-Mapped Tree Validation Summary"
      , subtitle = "height and DBH distribution comparison"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
      , strip.text = element_text(color = "black", size = 12)
      , strip.background = element_rect(fill = "gray88")
    )
```

### Detected Overstory ($TP$) Height Difference

Detected overstory tree ($TP$) height reliability. Where overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r regen-dta-sum-ht}
field_uas_comparison %>%
  dplyr::filter(field_uas_group == "true positive") %>%
  dplyr::group_by(plot_lab) %>% 
  dplyr::mutate(
    reorder_y = median(height_diff_pct, na.rm=T)
    , color_box = reorder_y<0 
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      plot_lab
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = height_diff_pct, y = reorder(plot_lab, reorder_y) )) + # height_diff_pct
    geom_vline(xintercept = 0, color = "gray22") +
    geom_boxplot(mapping = aes(fill = color_box), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    scale_fill_manual(values = c("steelblue", "coral")) +
    geom_point(color = "black", shape = ".") +
    scale_x_continuous(
      labels = scales::percent_format()
      , breaks = scales::extended_breaks(n=8)
      , limits = c(
        -max(field_uas_comparison$abs_height_diff_pct, na.rm = T)
        , max(field_uas_comparison$abs_height_diff_pct, na.rm = T)
      )
    ) +
    labs(
      fill = ""
      , y = ""
      , x = "Percent Difference in Height"
      , title = "Detected Overstory Height Difference"
      , subtitle = "by overstory plot"
      , caption = "-values = UAS<field | +values = UAS>field"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
    )
```

### Detected Overstory ($TP$) DBH Difference

Detected overstory tree ($TP$) DBH reliability. Where overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r regen-dta-sum-dbh}
field_uas_comparison %>%
  dplyr::filter(field_uas_group == "true positive") %>%
  dplyr::group_by(plot_lab) %>% 
  dplyr::mutate(
    reorder_y = median(dbh_diff_pct, na.rm=T)
    , color_box = reorder_y<0 
    , n_rows = dplyr::n()
    , plot_lab = paste0(
      plot_lab
      ," (n=", scales::comma(n_rows,accuracy=1),")"
    )
  ) %>% 
  ggplot(mapping = aes(x = dbh_diff_pct, y = reorder(plot_lab, reorder_y) )) + # dbh_diff_pct
    geom_vline(xintercept = 0, color = "gray22") +
    geom_boxplot(mapping = aes(fill = color_box), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    scale_fill_manual(values = c("steelblue", "coral")) +
    geom_point(color = "black", shape = ".") +
    scale_x_continuous(
      labels = scales::percent_format()
      , breaks = scales::extended_breaks(n=8)
      , limits = c(
        -max(field_uas_comparison$abs_dbh_diff_pct, na.rm = T)
        , max(field_uas_comparison$abs_dbh_diff_pct, na.rm = T)
      )
    ) +
    labs(
      fill = ""
      , y = ""
      , x = "Percent Difference in DBH"
      , title = "Detected Overstory DBH Difference"
      , subtitle = "by overstory plot"
      , caption = "-values = UAS<field | +values = UAS>field"
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
    )

```

### Detected Overstory ($TP$) Reliability

Overstory trees considered in this analysis had a tree height >= `r scales::comma(min_tree_height_m*3.28, accuracy = 0.1)` ft and a tree DBH >= `r scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)` in.

```{r regen-dta-sum-rmse, results='asis'}
dbh_f_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(field_dbh_in)
dbh_u_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(uas_dbh_in)
ht_f_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(field_tree_height_ft)
ht_u_temp = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% dplyr::pull(uas_tree_height_ft)

data.frame(
  mae_dbh = Metrics::mae(
    dbh_f_temp
    , dbh_u_temp
  )
  , mape_dbh = Metrics::mape(
    dbh_f_temp
    , dbh_u_temp
  )
  , rmse_dbh = Metrics::rmse(
    dbh_f_temp
    , dbh_u_temp
  )
    ## height
  , mae_height = Metrics::mae(
    ht_f_temp
    , ht_u_temp
  )
  , mape_height = Metrics::mape(
    ht_f_temp
    , ht_u_temp
  )
  , rmse_height = Metrics::rmse(
    ht_f_temp
    , ht_u_temp
  )
) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  tidyr::separate_wider_delim(cols = name, delim = "_", names = c("error", "metric")) %>% 
  tidyr::pivot_wider(names_from = error, values_from = value) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , n = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% nrow()
  ) %>% 
    kableExtra::kbl(
      caption = "Detected overstory tree height and DBH prediction performance"
      , col.names = c(
        " "
        , "Mean Abs. Error"
        , "Mean Abs. Percent Error"
        , "Root Mean Squared Error"
        , "N"
      )
      , digits = 2
    ) %>% 
    kableExtra::kable_styling()

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

### Detected Overstory ($TP$) Distribution Comparison

```{r regen-dta-sum-dist}
field_uas_comparison %>% 
  dplyr::filter(field_uas_group=="true positive") %>% 
  dplyr::select(c(
    tidyselect::ends_with("_dbh_in")
    , tidyselect::ends_with("_height_ft")
  )) %>% 
  dplyr::rename_with(
    .fn = function(x){dplyr::case_when(
      stringr::str_ends(x,"_dbh_in") ~ stringr::str_replace(x,"_dbh_in","_dbh")
      , stringr::str_ends(x,"_tree_height_ft") ~ stringr::str_replace(x,"_tree_height_ft","_height")
      , TRUE ~ x
    )}    
  ) %>% 
  tidyr::pivot_longer(dplyr::everything()) %>% 
  tidyr::separate_wider_delim(
    cols = name
    , delim = "_"
    , names = c("source", "metric")
  ) %>% 
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "dbh" ~ "DBH (in)"
      , metric == "height" ~ "Height (ft)"
    )
    , source = dplyr::case_when(
      source == "uas" ~ "UAS"
      , source == "field" ~ "Stem Map"
    )
    , n = field_uas_comparison %>% dplyr::filter(field_uas_group=="true positive") %>% nrow()
  ) %>% 
  ggplot(mapping = aes(x = value, y = source )) +
    geom_boxplot(mapping = aes(fill = source), alpha = 0.8, width = 0.7, outlier.shape = NA) +
    # geom_point(color = "black", shape = ".") +
    facet_grid(cols = vars(metric), scales = "free_x") +
    scale_fill_manual(values = c("seagreen", "slateblue")) +
    labs(
      fill = ""
      , y = ""
      , x = ""
      , title = "Detected Overstory Distribution Comparison"
      , subtitle = paste0(
        "tree height >= "
        , scales::comma(min_tree_height_m*3.28, accuracy = 0.1)
        , " ft and tree DBH >= "
        , scales::comma(min_tree_dbh_cm*0.394, accuracy = 0.1)
        , " in"
      )
    ) +
    theme_light() +
    theme(
      legend.position = "none"
      , axis.title.x = element_text(size=10, face = "bold")
      , axis.title.y = element_blank()
      , axis.text.y = element_text(color = "black",size=10, face = "bold", hjust = 0)
      , strip.text = element_text(color = "black", size = 12)
      , strip.background = element_rect(fill = "gray88")
    )

```

