---
title: "Black Hills Experimental Forest UAS Mission Data Summary"
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html')) 
  })
---

# Introduction

This project summarizes processed point cloud data created using structure from motion (SfM) photogrammetry methods on imagery collected from Unmanned Aircraft System (UAS) flights completed in June 2023. UAS imagery was collected using a DJI Phantom 4 Pro equipped with a 20 megapixel metal oxide semiconductor red-green-blue camera at a fixed 8.8 mm focal length. Flights followed a pre-programmed serpentine flight paths at an altitude of 120 m above ground level, using a nadir camera orientation, with 80% forward and side image overlap.The UAV imagery was processed in Agisoft Metashape using the routine outlined by [Tinkham and Swayze (2021)](https://scholar.google.com/scholar?cluster=11260597505702247290&hl=en&as_sdt=0,6) to produce an SfM point cloud for optimized tree detection. 

The SfM-derived point cloud was processed using the methods outlined by [Swayze and Tinkham (2022)](https://scholar.google.com/scholar?oi=bibs&hl=en&cluster=10655866445299954513). Specifically, [this script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/da5c8a13e9b2f2914b306201605e907643b3006d/src/point_cloud_processing.R) which is outlined in detail at it's [parent Github repository](https://georgewoolsey.github.io/point_cloud_tree_detection_ex/) was used to process the raw point cloud data.

## Objective

The objective of this analysis is to summarize forest stand characteristics using the UAS-derived tree list which includes tree location, height, and DBH measurements. Forest stands in this analysis include stands which were harvested in 2021 using single-tree and group-selection silvicultural prescriptions. 

The Forest Activity Tracking System (FACTS) [database](https://data.fs.usda.gov/geodata/edw/datasets.php?xmlKeyword) maintained by the U.S. Department of Agriculture, Forest Service (USFS) was used to delineate georeferenced boundaries of forest harvest activities.

# Setup
  
```{r, include=FALSE, warning=F, message=F}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results='hide'
  , fig.width = 10
  , fig.height = 7
)
```

## Load packages

Load all packages used in program.

```{r pkg-load}
# library
library(tidyverse) # the tidyverse
library(viridis) # viridis colors
library(scales) # work with number and plot scales
# spatial analysis
library(terra) # raster
library(sf) # simple features
# visualization
library(kableExtra)
library(patchwork) # ! better align plots in grid
library(mapview) # interactive html maps
# option to put satellite imagery as base layer of mapview maps
  mapview::mapviewOptions(
    homebutton = FALSE
    , basemaps = c("Esri.WorldImagery","OpenStreetMap")
  )
```

# Data Load and Setup

```{r dta-load}

```


```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```
