---
title: "Black Hills Experimental Forest UAS Mission Data Summary"
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html')) 
  })
---

# Introduction

This project summarizes processed point cloud data created using structure from motion (SfM) photogrammetry methods on imagery collected from Unmanned Aircraft System (UAS) flights completed in June 2023. UAS imagery was collected using a DJI Phantom 4 Pro equipped with a 20 megapixel metal oxide semiconductor red-green-blue camera at a fixed 8.8 mm focal length. Flights followed a pre-programmed serpentine flight paths at an altitude of 120 m above ground level, using a nadir camera orientation, with 80% forward and side image overlap.The UAV imagery was processed in Agisoft Metashape using the routine outlined by [Tinkham and Swayze (2021)](https://scholar.google.com/scholar?cluster=11260597505702247290&hl=en&as_sdt=0,6) to produce an SfM point cloud for optimized tree detection. 

The SfM-derived point cloud was processed using the methods outlined by [Swayze and Tinkham (2022)](https://scholar.google.com/scholar?oi=bibs&hl=en&cluster=10655866445299954513). Specifically, [this script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/da5c8a13e9b2f2914b306201605e907643b3006d/src/point_cloud_processing.R) which is outlined in detail at it's [parent Github repository](https://georgewoolsey.github.io/point_cloud_tree_detection_ex/) was used to process the raw point cloud data. UAS point cloud files (`laz` files) for this project were processed in two batches with the outputs combined using [this script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/6813f499e127b474dd91b480a993beb3d5225a11/src/point_cloud_processing_run_combine.R) from the same repository.

SfM techniques using UAS data reliably identify most tree locations and heights in open-canopied forests. However, their ability to characterize diameter at breast height (DBH) is limited to estimates of a fraction of trees within the area. SfM-derived DBH was estimated at the tree level following the methods outlined by [Tinkham et al. (2022)](https://scholar.google.com/scholar?cluster=14807895789640069059&hl=en&as_sdt=0,6).

## Objective

The objective of this analysis is to summarize forest stand characteristics using the UAS-derived tree list which includes tree location, height, and DBH measurements. Forest stands in this analysis include stands which were harvested in 2021 using single-tree and group-selection silvicultural prescriptions. 

The Forest Activity Tracking System (FACTS) [database](https://data.fs.usda.gov/geodata/edw/datasets.php?xmlKeyword) maintained by the U.S. Department of Agriculture, Forest Service (USFS) was used to delineate georeferenced boundaries of forest harvest activities.

# Setup
  
```{r, include=FALSE, warning=F, message=F}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results='hide'
  , fig.width = 10
  , fig.height = 7
)
```

## Load packages

Load all packages used in program.

```{r pkg-load}
# library
library(tidyverse) # the tidyverse
library(viridis) # viridis colors
library(scales) # work with number and plot scales
library(latex2exp) # math formulas text
# spatial analysis
library(terra) # raster
library(sf) # simple features
library(stars) # stars raster
# visualization
library(kableExtra)
library(patchwork) # ! better align plots in grid
library(mapview) # interactive html maps
library(leafpop) # interactive html maps
library(ggnewscale) # add another scale to ggplot
library(ggtext) # customize plot text colors
library(pdftools) # combine pdf files
# modelling
library(brms) # bayesian modelling using STAN engine

# option to put satellite imagery as base layer of mapview maps
  mapview::mapviewOptions(
    homebutton = FALSE
    , basemaps = c("Esri.WorldImagery","OpenStreetMap")
  )
```

# Data Load and Setup

```{r dta-load, results='hide'}
# where is the processed data from point_cloud_processing_run_combine.R ?
input_dir = "../data/point_cloud_processing_BHEF_202306_combined"
# set output directory
delivery_dir = "../data/delivery"
dir.create(delivery_dir, showWarnings = FALSE)

### read data from point_cloud_processing
#rasters
dtm_rast = terra::rast(paste0(input_dir, "/dtm_1m.tif"))
chm_rast = terra::rast(paste0(input_dir, "/chm_0.25m.tif"))
# crowns = terra::rast(paste0(input_dir, "/top_down_detected_tree_crowns.tif"))

# las extent
las_ctg_dta = sf::st_read(paste0(input_dir, "/raw_las_ctg_info.gpkg"))

# vectors
# crown polygons
  crowns_sf_with_dbh = sf::st_read(paste0(input_dir, "/final_detected_crowns.gpkg")) %>% 
    dplyr::select(-c(basal_area_ft2)) %>% 
    dplyr::rename(
      reg_est_lower_dbh_cm = reg_est_dbh_cm_lower
      , reg_est_upper_dbh_cm = reg_est_dbh_cm_upper
    )

# tree top points
treetops_sf_with_dbh = sf::st_read(paste0(input_dir, "/final_detected_tree_tops.gpkg")) %>% 
  dplyr::select(-c(basal_area_ft2)) %>% 
  dplyr::rename(
    reg_est_lower_dbh_cm = reg_est_dbh_cm_lower
    , reg_est_upper_dbh_cm = reg_est_dbh_cm_upper
  )

# height range for use in estimation
height_range = dplyr::tibble(
  tree_height_m = seq(
    from = 0
    , to = round(max(treetops_sf_with_dbh$tree_height_m)*1.05,0)
    , by = 0.1 # by 0.1 m increments
  )
)

# set crs
proj_crs = sf::st_crs(las_ctg_dta)

# read data from FACTS
# harvests
  harvests = sf::st_read("../data/bhef_harvests.gpkg") %>% 
    dplyr::filter(
      year_id == 2021
      # year_id >= year(Sys.time()) - 15 
      & !(treatment_type_grp %in% c("Improvement/Liberation Cut", "Other", "Sanitation Cut"))
    ) %>% 
    sf::st_transform(proj_crs)
# bhef bounds
  bhef_boundary = sf::st_read("../data/bhef_boundary.gpkg") %>% 
    sf::st_transform(proj_crs)
```

## Load orthomosaic rasters

```{r ortho-ld, results='hide'}
# read list of orthos
ortho_list_temp = list.files("../data/ortho/", pattern = ".*\\.(tif|tiff)$", full.names = T)[] %>% 
    purrr::map(function(x){terra::rast(x)})

# ortho_list_temp[[1]] %>%
#   terra::aggregate(20) %>%
#   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "hist", colNA = "transparent")

####### ensure the resolution of the rasters matches 
  # terra::res(ortho_list_temp[[1]])
  
  ## function
  change_res_fn = function(r, my_res=1){
    r2 = r
    terra::res(r2) = my_res
    r2 = terra::resample(r, r2)
    return(r2)
  }
  ## apply the function
  ortho_list_temp = 1:length(ortho_list_temp) %>% 
    purrr::map(function(x){change_res_fn(ortho_list_temp[[x]], my_res=0.5)})
  
  # terra::res(ortho_list_temp[[1]])
  
  # ortho_list_temp[[1]] %>%
  #   terra::aggregate(2) %>%
  #   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "hist", colNA = "transparent")

######## mosaic the raster list 
  ortho_rast = terra::mosaic(
    terra::sprc(ortho_list_temp)
    , fun = "min" # min only thing that works
  ) 

  # ortho_rast %>%
  #   # terra::aggregate(2) %>%
  #   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "lin", colNA = "transparent")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

# Local DBH to Height Model

## Random Forest Model

The UAS point cloud [processing script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/da5c8a13e9b2f2914b306201605e907643b3006d/src/point_cloud_processing.R) used a random forest model to estimate tree DBH based on the sample of DBH values extracted from the point cloud (outlined in detail [here](https://georgewoolsey.github.io/point_cloud_tree_detection_ex/#Model_to_estimate_missing_DBHs). This random forest model included tree height (m), crown area (m^2^), minimum crown height (m), and local tree competition metrics to estimate missing DBH values. Using methods outlined by [Tinkham et al. (2022)](https://scholar.google.com/scholar?cluster=14807895789640069059&hl=en&as_sdt=0,6), local competition metrics included the distance to the nearest neighbor, trees ha^−1^ within a 5 m radius, and the relative tree height within a 5 m radius. A 5 m radius was selected as it represents a distance slightly less than the point at which two mature ponderosa pine trees in the region would begin to have interlocking crowns, representing a proxy for direct competition. The relative tree height within a 5 m radius was estimated using:

$$
\textrm{Relative Height} = \frac{\textrm{Height}}{\textrm{Height}_\textrm{max}} \times 100
$$

, where $\textrm{Height}$ is the height of the subject tree and $\textrm{Height}_\textrm{max}$ is the height of the tallest tree within a 5 m radius of the subject tree.

### Training vs Non-Training DBH

```{r plt-rf}
### plot
plt_rf = treetops_sf_with_dbh %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_training_data==T) %>% 
  dplyr::bind_rows(
    treetops_sf_with_dbh %>% 
    sf::st_drop_geometry() %>% 
    dplyr::filter(is_training_data==F) %>% 
    dplyr::slice_sample(prop = 0.15)
  ) %>% 
  ggplot(
    mapping = aes(
      x=tree_height_m, y = dbh_cm
      , color = is_training_data
      , fill = is_training_data
      , group = is_training_data
    )
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.7
    , color = "transparent"
    , shape = 21
    , show.legend = F
  ) + 
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , lwd = 1.5
    , show.legend = T
  ) +
  scale_color_manual(values = c("gray","firebrick")) +
  scale_fill_manual(values = c("gray","firebrick")) +
  scale_y_continuous(limits = c(0,max(treetops_sf_with_dbh$dbh_cm)*1.02)) +
  scale_x_continuous(limits = c(0,max(treetops_sf_with_dbh$tree_height_m)*1.02)) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Random Forest Model"
    , subtitle = "SfM-derived tree height and DBH relationship"
    , color = "Training Data"
    , fill = "Training Data"
  ) +
  theme_light() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, lwd = 6, size = 6, alpha = 1))
  )
# plot
plt_rf
```

## Linear Model

[Tinkham et al. (2022)](https://scholar.google.com/scholar?cluster=14807895789640069059&hl=en&as_sdt=0,6) found the simple linear form resulted in the best estimates for both tree and stand level summarizations:

\begin{align*}
DBH_{i}  &\sim {\sf Gamma} \bigl(g(\beta_0, \beta_1, Height_{i}), shape \bigr) \\
g(\beta_0, \beta_1, Height_{i}) &= {\exp \bigl(\beta_0 + \beta_1 \cdot Height_{i}\bigr)} \\
\beta_{0}  &\sim {\sf Normal} (0,10) \\ 
\beta_{1}  &\sim {\sf Normal} (0,10) \\ 
shape  &\sim {\sf Gamma} (0.01,0.01)
\end{align*}

, where $i$ is the individual tree estimate from UAS SfM data. The $\sf Gamma$ response distribution represents $DBH_{i}$ as strictly non-negative.

```{r mod-lin, results='hide'}
# population model with no random effects (i.e. no group-level variation)
  # Gamma distribution for strictly positive response variable dbh
  mod_lin = brms::brm(
    formula = dbh_cm ~ 1 + tree_height_m
    , data = treetops_sf_with_dbh %>% 
        sf::st_drop_geometry() %>% 
        dplyr::filter(is_training_data==T)
    , family = brms::brmsfamily("Gamma", link = "identity")
    , prior = c(prior(gamma(0.01, 0.01), class = shape))
    , iter = 3000
    , file = "../data/mod_lin"
  )
  # plot(mod_lin)
  # summary(mod_lin)
  # plot(brms::conditional_effects(mod_lin), points = T)

  # compute and save the WAIC information for the next three models
  mod_lin = brms::add_criterion(mod_lin, c("loo", "waic"))
  
  # predict values over range
  pred_mod_lin = predict(
      mod_lin
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character())
  
  # attach predicted values to original data
  treetops_sf_with_dbh = treetops_sf_with_dbh %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character()) %>% 
    dplyr::left_join(
       pred_mod_lin %>% 
        dplyr::select(tree_height_m_tnth, estimate)
      , by = dplyr::join_by(tree_height_m_tnth)
    ) %>% 
    dplyr::mutate(
      mod_lin_est_dbh_cm = ifelse(is_training_data == T, dbh_cm, estimate)
    ) %>% 
    dplyr::select(-c(estimate, tree_height_m_tnth))
  
  gc()
```

### Model Parameter Estimates

Simple linear form:

$$
DBH_i = \beta_0 + \beta_1 \cdot Height_i
$$

, where $i$ is the individual tree estimate from UAS SfM data

```{r mod-lin-params, results='asis'}
# parameter estimates
mod_lin_sum_temp = brms::posterior_summary(mod_lin) %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>% 
  dplyr::filter(stringr::str_starts(parameter, "b_")) %>% # | parameter == "shape") %>% 
  dplyr::rename_with(tolower)
# table
mod_lin_sum_temp %>% 
    kableExtra::kbl(digits = 4) %>% 
    kableExtra::kable_styling()
```

### SfM-derived Height to DBH Allometry

```{r mod-lin-pred}
# plot predictions with data
ggplot(
    data = pred_mod_lin, mapping = aes(x = tree_height_m)
  ) +
  geom_point(
    data = treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(is_training_data==T)
    , mapping = aes(y = dbh_cm)
    , color = "firebrick"
  ) +
  geom_ribbon(
    mapping = aes(ymin = lower_b, ymax = upper_b)
    , fill = "gray88"
    , alpha = 0.5
  ) +
  geom_line(
    aes(y = estimate)
    , color = "gray33"
    , lwd = 1
  ) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Local height to DBH allometry from SfM-extracted DBH samples"
    , subtitle = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_lin_sum_temp %>% dplyr::filter(parameter=="b_Intercept") %>% dplyr::pull(estimate) %>% round(3)
      , " + "
      , mod_lin_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
    ))
  ) +
  theme_light() +
  theme(legend.position = "none")
```

### Training vs Non-Training DBH

```{r plt-lin}
### plot
plt_lin = treetops_sf_with_dbh %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_training_data==T) %>% 
  dplyr::bind_rows(
    treetops_sf_with_dbh %>% 
    sf::st_drop_geometry() %>% 
    dplyr::filter(is_training_data==F) %>% 
    dplyr::slice_sample(prop = 0.15)
  ) %>% 
  ggplot(
    mapping = aes(
      x=tree_height_m, y = mod_lin_est_dbh_cm
      , color = is_training_data
      , fill = is_training_data
      , group = is_training_data
    )
  ) +
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , lwd = 1.5
    , show.legend = T
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.7
    , color = "transparent"
    , shape = 21
    , show.legend = F
  ) + 
  annotate(
    "text", x = 0, y = max(treetops_sf_with_dbh$dbh_cm)
    , label = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_lin_sum_temp %>% dplyr::filter(parameter=="b_Intercept") %>% dplyr::pull(estimate) %>% round(3)
      , " + "
      , mod_lin_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
    ))
    , parse = T
    , hjust = 0
  ) +
  scale_color_manual(values = c("gray","firebrick")) +
  scale_fill_manual(values = c("gray","firebrick")) +
  scale_y_continuous(limits = c(0,max(treetops_sf_with_dbh$dbh_cm)*1.02)) +
  scale_x_continuous(limits = c(0,max(treetops_sf_with_dbh$tree_height_m)*1.02)) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Linear Model"
    , subtitle = "SfM-derived tree height and DBH relationship"
    , color = "Training Data"
    , fill = "Training Data"
  ) +
  theme_light() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, lwd = 6, size = 6, alpha = 1))
  )
# plot
plt_lin
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Quadratic Model

A quadratic linear model with the inclusion of $Height_{i}^{2}$ in the model allows us to account for a nonlinear (e.g. parabolic) relationship between DBH and tree height.

\begin{align*}
DBH_{i}  &\sim {\sf Gamma} \bigl(g(\beta_0, \beta_1, \beta_2, Height_{i}), shape \bigr) \\
g(\beta_0, \beta_1, \beta_2, Height_{i}) &= {\exp \bigl(\beta_0 + \beta_1 \cdot Height_{i} + \beta_2 \cdot Height_{i}^{2}\bigr)} \\
\beta_{0}  &\sim {\sf Normal} (0,10) \\ 
\beta_{1}  &\sim {\sf Normal} (0,10) \\ 
\beta_{2}  &\sim {\sf Normal} (0,10) \\ 
shape  &\sim {\sf Gamma} (0.01,0.01)
\end{align*}

, where $i$ is the individual tree estimate from UAS SfM data. The $\sf Gamma$ response distribution represents $DBH_{i}$ as strictly non-negative.

```{r mod-quad, results='hide'}
# population model with no random effects (i.e. no group-level variation)
  # Gamma distribution for strictly positive response variable dbh
  mod_quad = brms::brm(
    formula = dbh_cm ~ 1 + tree_height_m + I(tree_height_m^2)
    , data = treetops_sf_with_dbh %>% 
        sf::st_drop_geometry() %>% 
        dplyr::filter(is_training_data==T)
    , family = brms::brmsfamily("Gamma", link = "log")
    , prior = c(prior(gamma(0.01, 0.01), class = shape))
    , iter = 3000
    , file = "../data/mod_quad"
  )
  # plot(mod_quad)
  # summary(mod_quad)
  # plot(brms::conditional_effects(mod_quad), points = T)

  # compute and save the WAIC information for the next three models
  mod_quad = brms::add_criterion(mod_quad, c("loo", "waic"))
  
  # predict values over range
  pred_mod_quad = predict(
      mod_quad
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character())
  
  # attach predicted values to original data
  treetops_sf_with_dbh = treetops_sf_with_dbh %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character()) %>% 
    dplyr::left_join(
       pred_mod_quad %>% 
        dplyr::select(tree_height_m_tnth, estimate)
      , by = dplyr::join_by(tree_height_m_tnth)
    ) %>% 
    dplyr::mutate(
      mod_quad_est_dbh_cm = ifelse(is_training_data == T, dbh_cm, estimate)
    ) %>% 
    dplyr::select(-c(estimate, tree_height_m_tnth))
  
  gc()
```

### Model Parameter Estimates

Quadratic linear model form:

$$
DBH_{i} = \beta_0 + \beta_1 \cdot Height_{i} + \beta_2 \cdot Height_{i}^{2}
$$

, where $i$ is the individual tree estimate from UAS SfM data

```{r mod-quad-params, results='asis'}
# parameter estimates
mod_quad_sum_temp = brms::posterior_summary(mod_quad) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>% 
  dplyr::filter(stringr::str_starts(parameter, "b_")) %>% # | parameter == "shape") %>% 
  dplyr::rename_with(tolower) %>% 
  dplyr::mutate(
    dplyr::across(.cols = tidyselect::where(is.numeric), .fns = exp)
  )
# table
mod_quad_sum_temp %>% 
    kableExtra::kbl(digits = 4) %>% 
    kableExtra::kable_styling()
```

### SfM-derived Height to DBH Allometry

```{r mod-quad-pred}
# predict and put estimates in a data frame
  predict(
      mod_quad
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
# plot predictions with data
ggplot(
    mapping = aes(x = tree_height_m)
  ) +
  geom_point(
    data = treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(is_training_data==T)
    , mapping = aes(y = dbh_cm)
    , color = "firebrick"
  ) +
  geom_ribbon(
    mapping = aes(ymin = lower_b, ymax = upper_b)
    , fill = "gray88"
    , alpha = 0.5
  ) +
  geom_line(
    aes(y = estimate)
    , color = "gray33"
    , lwd = 1
  ) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Local height to DBH allometry from SfM-extracted DBH samples"
    , subtitle = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_Intercept") %>% dplyr::pull(estimate) %>% round(3)
      , " + "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
      , " + "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_Itree_height_mE2") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^2$"
    ))
  ) +
  theme_light() +
  theme(legend.position = "none")
```

### Training vs Non-Training DBH

```{r plt-quad}
### plot
plt_quad = treetops_sf_with_dbh %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_training_data==T) %>% 
  dplyr::bind_rows(
    treetops_sf_with_dbh %>% 
    sf::st_drop_geometry() %>% 
    dplyr::filter(is_training_data==F) %>% 
    dplyr::slice_sample(prop = 0.05)
  ) %>% 
  ggplot(
    mapping = aes(
      x=tree_height_m, y = mod_quad_est_dbh_cm
      , color = is_training_data
      , fill = is_training_data
      , group = is_training_data
    )
  ) +
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , lwd = 1.5
    , show.legend = T
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.7
    , color = "transparent"
    , shape = 21
    , show.legend = F
  ) + 
  annotate(
    "text", x = 0, y = max(treetops_sf_with_dbh$dbh_cm)
    , label = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_Intercept") %>% dplyr::pull(estimate) %>% round(3)
      , " + "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
      , " + "
      , mod_quad_sum_temp %>% dplyr::filter(parameter=="b_Itree_height_mE2") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^2$"
    ))
    , parse = T
    , hjust = 0
  ) +
  scale_color_manual(values = c("gray","firebrick")) +
  scale_fill_manual(values = c("gray","firebrick")) +
  scale_y_continuous(limits = c(0,max(treetops_sf_with_dbh$dbh_cm)*1.02)) +
  scale_x_continuous(limits = c(0,max(treetops_sf_with_dbh$tree_height_m)*1.02)) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Quadratic Model"
    , subtitle = "SfM-derived tree height and DBH relationship"
    , color = "Training Data"
    , fill = "Training Data"
  ) +
  theme_light() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, lwd = 6, size = 6, alpha = 1))
  )
# plot
plt_quad
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## No Intercept Quadratic Model

A quadratic linear model with the inclusion of $Height_{i}^{2}$ in the model allows us to account for a nonlinear (e.g. parabolic) relationship between DBH and tree height. Forcing the intercept to $-1$ allows for the model to better account for the minimum tree height threshold set at `2 m` in the point cloud processing workflow.

\begin{align*}
DBH_{i}  &\sim {\sf Gamma} \bigl(g(\beta_1, \beta_2, Height_{i}), shape \bigr) \\
g(\beta_1, \beta_2, Height_{i}) &= {\exp \bigl(-1 + \beta_1 \cdot Height_{i} + \beta_2 \cdot Height_{i}^{2}\bigr)} \\
\beta_{1}  &\sim {\sf Normal} (0,10) \\ 
\beta_{2}  &\sim {\sf Normal} (0,10) \\ 
shape  &\sim {\sf Gamma} (0.01,0.01)
\end{align*}

, where $i$ is the individual tree estimate from UAS SfM data. The $\sf Gamma$ response distribution represents $DBH_{i}$ as strictly non-negative.

```{r mod-quad-noint, results='hide'}
# population model with no random effects (i.e. no group-level variation)
  # Gamma distribution for strictly positive response variable dbh
  mod_quad_noint = brms::brm(
    formula = dbh_cm ~ -1 + tree_height_m + I(tree_height_m^2)
    , data = treetops_sf_with_dbh %>% 
        sf::st_drop_geometry() %>% 
        dplyr::filter(is_training_data==T)
    , family = brms::brmsfamily("Gamma", link = "log")
    , prior = c(prior(gamma(0.01, 0.01), class = shape))
    , iter = 3000
    , file = "../data/mod_quad_noint"
  )
  # plot(mod_quad_noint)
  # summary(mod_quad_noint)
  # plot(brms::conditional_effects(mod_quad_noint), points = T)

  # compute and save the WAIC information for the next three models
  mod_quad_noint = brms::add_criterion(mod_quad_noint, c("loo", "waic"))
  
  # predict values over range
  pred_mod_quad_noint = predict(
      mod_quad_noint
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character())
  
  # attach predicted values to original data
  treetops_sf_with_dbh = treetops_sf_with_dbh %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character()) %>% 
    dplyr::left_join(
       pred_mod_quad_noint %>% 
        dplyr::select(tree_height_m_tnth, estimate)
      , by = dplyr::join_by(tree_height_m_tnth)
    ) %>% 
    dplyr::mutate(
      mod_quad_noint_est_dbh_cm = ifelse(is_training_data == T, dbh_cm, estimate)
    ) %>% 
    dplyr::select(-c(estimate, tree_height_m_tnth))
  
  gc()
```

### Model Parameter Estimates

No intercept quadratic linear model form:

$$
DBH_{i} = -1 + \beta_1 \cdot Height_{i} + \beta_2 \cdot Height_{i}^{2}
$$

, where $i$ is the individual tree estimate from UAS SfM data

```{r mod-quad-noint-params, results='asis'}
# parameter estimates
mod_quad_noint_sum_temp = brms::posterior_summary(mod_quad_noint) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>% 
  dplyr::filter(stringr::str_starts(parameter, "b_")) %>% # | parameter == "shape") %>% 
  dplyr::rename_with(tolower) %>% 
  dplyr::mutate(
    dplyr::across(.cols = tidyselect::where(is.numeric), .fns = exp)
  )
# table
mod_quad_noint_sum_temp %>% 
    kableExtra::kbl(digits = 4) %>% 
    kableExtra::kable_styling()
```

### SfM-derived Height to DBH Allometry

```{r mod-quad-noint-pred}
# predict and put estimates in a data frame
  predict(
      mod_quad_noint
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
# plot predictions with data
ggplot(
    mapping = aes(x = tree_height_m)
  ) +
  geom_point(
    data = treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(is_training_data==T)
    , mapping = aes(y = dbh_cm)
    , color = "firebrick"
  ) +
  geom_ribbon(
    mapping = aes(ymin = lower_b, ymax = upper_b)
    , fill = "gray88"
    , alpha = 0.5
  ) +
  geom_line(
    aes(y = estimate)
    , color = "gray33"
    , lwd = 1
  ) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Local height to DBH allometry from SfM-extracted DBH samples"
    , subtitle = latex2exp::TeX(paste0(
      "$DBH = -1"
      , " + "
      , mod_quad_noint_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
      , " + "
      , mod_quad_noint_sum_temp %>% dplyr::filter(parameter=="b_Itree_height_mE2") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^2$"
    ))
  ) +
  theme_light() +
  theme(legend.position = "none")
```

### Training vs Non-Training DBH

```{r plt-quad-noint}
### plot
plt_quad_noint = treetops_sf_with_dbh %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_training_data==T) %>% 
  dplyr::bind_rows(
    treetops_sf_with_dbh %>% 
    sf::st_drop_geometry() %>% 
    dplyr::filter(is_training_data==F) %>% 
    dplyr::slice_sample(prop = 0.05)
  ) %>% 
  ggplot(
    mapping = aes(
      x=tree_height_m, y = mod_quad_noint_est_dbh_cm
      , color = is_training_data
      , fill = is_training_data
      , group = is_training_data
    )
  ) +
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , lwd = 1.5
    , show.legend = T
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.7
    , color = "transparent"
    , shape = 21
    , show.legend = F
  ) + 
  annotate(
    "text", x = 0, y = max(treetops_sf_with_dbh$dbh_cm)
    , label = latex2exp::TeX(paste0(
      "$DBH = -1"
      , " + "
      , mod_quad_noint_sum_temp %>% dplyr::filter(parameter=="b_tree_height_m") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height"
      , " + "
      , mod_quad_noint_sum_temp %>% dplyr::filter(parameter=="b_Itree_height_mE2") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^2$"
    ))
    , parse = T
    , hjust = 0
  ) +
  scale_color_manual(values = c("gray","firebrick")) +
  scale_fill_manual(values = c("gray","firebrick")) +
  scale_y_continuous(limits = c(0,max(treetops_sf_with_dbh$dbh_cm)*1.02)) +
  scale_x_continuous(limits = c(0,max(treetops_sf_with_dbh$tree_height_m)*1.02)) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "No Intercept Quadratic Model"
    , subtitle = "SfM-derived tree height and DBH relationship"
    , color = "Training Data"
    , fill = "Training Data"
  ) +
  theme_light() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, lwd = 6, size = 6, alpha = 1))
  )
# plot
plt_quad_noint
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Nonlinear Model

A power-law function model allows us to account for a nonlinear relationship between DBH and tree height:

\begin{align*}
DBH_{i}  &\sim {\sf Gamma} \bigl(g(\alpha, \kappa, Height_{i}), shape \bigr) \\
g(\alpha, \kappa, Height_{i}) &= {\exp \bigl(\alpha \cdot Height_{i}^{\kappa}\bigr)} \\
\alpha  &\sim {\sf Normal} (0,10) \\ 
\kappa  &\sim {\sf Normal} (0,10) \\ 
shape  &\sim {\sf Gamma} (0.01,0.01)
\end{align*}

, where $i$ is the individual tree estimate from UAS SfM data. The $\sf Gamma$ response distribution represents $DBH_{i}$ as strictly non-negative.

```{r mod-pwr, results='hide'}
# population model with no random effects (i.e. no group-level variation)
  # Gamma distribution for strictly positive response variable dbh
  mod_pwr = brms::brm(
    formula = brms::bf(
      formula = dbh_cm ~ a * (tree_height_m^k)
      , a + k ~ 1
      , nl = TRUE # !! specify non-linear
    )
    , data = treetops_sf_with_dbh %>% 
        sf::st_drop_geometry() %>% 
        dplyr::filter(is_training_data==T)
    , family = brms::brmsfamily("Gamma", link = "log")
    , prior = c(prior(gamma(0.01, 0.01), class = shape))
    , iter = 3000
    , file = "../data/mod_pwr"
  )
  # plot(mod_pwr)
  # summary(mod_pwr)
  # plot(brms::conditional_effects(mod_pwr), points = T)

  # compute and save the WAIC information for the next three models
  mod_pwr = brms::add_criterion(mod_pwr, c("loo", "waic"))
  
  # predict values over range
  pred_mod_pwr = predict(
      mod_pwr
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character())
  
  # attach predicted values to original data
  treetops_sf_with_dbh = treetops_sf_with_dbh %>% 
    dplyr::mutate(tree_height_m_tnth = round(tree_height_m,1) %>% as.character()) %>% 
    dplyr::left_join(
       pred_mod_pwr %>% 
        dplyr::select(tree_height_m_tnth, estimate)
      , by = dplyr::join_by(tree_height_m_tnth)
    ) %>% 
    dplyr::mutate(
      mod_pwr_est_dbh_cm = ifelse(is_training_data == T, dbh_cm, estimate)
    ) %>% 
    dplyr::select(-c(estimate, tree_height_m_tnth))
  
  gc()
```

### Model Parameter Estimates

Power-law function model form:

$$
DBH_{i} = \alpha \cdot Height_{i}^{\kappa}
$$

, where $i$ is the individual tree estimate from UAS SfM data

```{r mod-pwr-params, results='asis'}
# parameter estimates
mod_pwr_sum_temp = brms::posterior_summary(mod_pwr) %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "parameter") %>% 
  dplyr::filter(stringr::str_starts(parameter, "b_")) %>% # | parameter == "shape") %>% 
  dplyr::rename_with(tolower) %>% 
  dplyr::mutate(
    parameter = stringr::str_remove_all(tolower(parameter), "_intercept")
    , dplyr::across(.cols = tidyselect::where(is.numeric), .fns = exp)
  )
# table
mod_pwr_sum_temp %>% 
    kableExtra::kbl(digits = 4) %>% 
    kableExtra::kable_styling()
```

### SfM-derived Height to DBH Allometry

```{r mod-pwr-pred}
# predict and put estimates in a data frame
  predict(
      mod_pwr
      , newdata = height_range
      , probs = c(.05, .95)
    ) %>%
    dplyr::as_tibble() %>%
    dplyr::rename(
      lower_b = 3, upper_b = 4
    ) %>% 
    dplyr::rename_with(tolower) %>% 
    dplyr::select(-c(est.error)) %>% 
    dplyr::bind_cols(height_range) %>% 
# plot predictions with data
ggplot(
    mapping = aes(x = tree_height_m)
  ) +
  geom_point(
    data = treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::filter(is_training_data==T)
    , mapping = aes(y = dbh_cm)
    , color = "firebrick"
  ) +
  geom_ribbon(
    mapping = aes(ymin = lower_b, ymax = upper_b)
    , fill = "gray88"
    , alpha = 0.5
  ) +
  geom_line(
    aes(y = estimate)
    , color = "gray33"
    , lwd = 1
  ) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Local height to DBH allometry from SfM-extracted DBH samples"
    , subtitle = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_pwr_sum_temp %>% dplyr::filter(parameter=="b_a") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^{"
      , mod_pwr_sum_temp %>% dplyr::filter(parameter=="b_k") %>% dplyr::pull(estimate) %>% round(3)
      , "}$"
    ))
  ) +
  theme_light() +
  theme(legend.position = "none")
```

### Training vs Non-Training DBH

```{r plt-pwr}
### plot
plt_pwr = treetops_sf_with_dbh %>% 
  sf::st_drop_geometry() %>% 
  dplyr::filter(is_training_data==T) %>% 
  dplyr::bind_rows(
    treetops_sf_with_dbh %>% 
    sf::st_drop_geometry() %>% 
    dplyr::filter(is_training_data==F) %>% 
    dplyr::slice_sample(prop = 0.05)
  ) %>% 
  ggplot(
    mapping = aes(
      x=tree_height_m, y = mod_pwr_est_dbh_cm
      , color = is_training_data
      , fill = is_training_data
      , group = is_training_data
    )
  ) +
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , lwd = 1.5
    , show.legend = T
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.7
    , color = "transparent"
    , shape = 21
    , show.legend = F
  ) + 
  annotate(
    "text", x = 0, y = max(treetops_sf_with_dbh$dbh_cm)
    , label = latex2exp::TeX(paste0(
      "$DBH = "
      , mod_pwr_sum_temp %>% dplyr::filter(parameter=="b_a") %>% dplyr::pull(estimate) %>% round(3)
      , "\\cdot Height^{"
      , mod_pwr_sum_temp %>% dplyr::filter(parameter=="b_k") %>% dplyr::pull(estimate) %>% round(3)
      , "}$"
    ))
    , parse = T
    , hjust = 0
  ) +
  scale_color_manual(values = c("gray","firebrick")) +
  scale_fill_manual(values = c("gray","firebrick")) +
  scale_y_continuous(limits = c(0,max(treetops_sf_with_dbh$dbh_cm)*1.02)) +
  scale_x_continuous(limits = c(0,max(treetops_sf_with_dbh$tree_height_m)*1.02)) +
  labs(
    y = "DBH (cm)"
    , x = "Tree Ht. (m)"
    , title = "Nonlinear Model"
    , subtitle = "SfM-derived tree height and DBH relationship"
    , color = "Training Data"
    , fill = "Training Data"
  ) +
  theme_light() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
  ) +
  guides(
    color = guide_legend(override.aes = list(shape = 15, lwd = 6, size = 6, alpha = 1))
  )
# plot
plt_pwr
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Model Comparison

The pointwise log-likelihood can be used, among others, to calculate the leave-one-out cross-validation (LOO; [Vehtari, Gelman, and Gabry 2015](https://scholar.google.com/scholar?cluster=13592897121719449855&oi=gsb&hl=en&as_sdt=0,6)) allowing for comparison of different models applied to the same data (lower LOOs indicate better model fit).

```{r mod-comp, results='asis'}
brms::loo_compare(mod_lin, mod_quad, mod_quad_noint, mod_pwr, criterion = "loo") %>% 
  kableExtra::kbl(digits = 3) %>% 
  kableExtra::kable_styling()
# # can also check model weights
# brms::model_weights(mod_lin, mod_quad, mod_quad_noint, mod_pwr, weights = "loo") %>% 
#   round(3)
```

Based on these results, the quadratic model $DBH_i = \beta_0 + \beta_1 \cdot Height_i + \beta_2 \cdot Height_{i}^{2}$ has the best fit among models tested. DBH estimates based on this model for trees where DBH could not be extracted from the UAS SfM data will be used in the following analysis and stand-level summaries

## Update DBH to best model
 
```{r dbh-update}
# update treetops
treetops_sf_with_dbh = treetops_sf_with_dbh %>%
  dplyr::mutate(
    mod_rf_est_dbh_cm = dbh_cm
    , dbh_cm = mod_quad_est_dbh_cm
    , dbh_m = dbh_cm/100
    , radius_m = dbh_m/2
    , basal_area_m2 = pi * (radius_m)^2
  ) %>% 
  # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.28
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    )

##########
# update crowns
crowns_sf_with_dbh = crowns_sf_with_dbh %>% 
  dplyr::select(-c(
    dbh_cm
    , dbh_m
    , radius_m
    , basal_area_m2
  )) %>% 
  dplyr::inner_join(
    treetops_sf_with_dbh %>% 
      sf::st_drop_geometry() %>% 
      dplyr::select(
        treeID
        , dbh_cm
        , dbh_m
        , radius_m
        , basal_area_m2
        , (tidyselect::starts_with("mod_") & tidyselect::ends_with("_cm"))
      )
    , by = dplyr::join_by(treeID)
  ) %>% 
  # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.28
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    )
    
```


# Combine harvest units with tree locations

Spatially combine harvest units with tree locations. One tree can be located in multiple harvest units if the harvest units spatially overlap; I'll allow it. Trees are matched to a harvest unit based on the location of the tree top which may or may not align with the location of the tree bole at DBH.

```{r harvest-trees}
#################################################################################
#################################################################################
# Join tree tops with forest stands
#################################################################################
#################################################################################
harvests_trees = harvests %>%
  dplyr::mutate(
    stand_area_m2 = sf::st_area(.) %>% as.numeric()
    , stand_area_ha = stand_area_m2/10000
  ) %>% 
  sf::st_intersection(las_ctg_dta) %>% 
  dplyr::mutate(
    intrsct_stand_area_m2 = sf::st_area(.) %>% as.numeric()
  ) %>% 
  dplyr::filter(round(intrsct_stand_area_m2, 0) == round(stand_area_m2, 0)) %>% 
  dplyr::select(
    suid, forest_commonname, admin_region_code, activity_name
    , treatment_type, treatment_type_grp, date_compl, year_id
    , stand_area_m2, stand_area_ha
  ) %>% 
  sf::st_intersection(treetops_sf_with_dbh)
# filter harvests
harvests = harvests %>% 
  dplyr::filter(suid %in% unique(harvests_trees$suid))

```

# Exploratory Analysis

High-level summaries of the data.

## Study Location Map
```{r stats-temp, include=FALSE, eval=TRUE}
ac_temp = las_ctg_dta %>% sf::st_area() %>% as.numeric() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
ha_temp = las_ctg_dta %>% sf::st_area() %>% as.numeric() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
min_ac_temp = harvests %>% sf::st_area() %>% as.numeric() %>% min() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
min_ha_temp = harvests %>% sf::st_area() %>% as.numeric() %>% min() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
max_ac_temp = harvests %>% sf::st_area() %>% as.numeric() %>% max() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
max_ha_temp = harvests %>% sf::st_area() %>% as.numeric() %>% max() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
```

Map of the BHEF, 2021 harvest boundaries, and 2023-06 UAS flight boundaries. The UAS flights covered `r ac_temp` acres (`r ha_temp` ha). There are `r nrow(harvests)` stands harvested in 2021 that are entirely within the UAS flight boundaries. The harvested stands range in size from `r min_ac_temp` acres (`r min_ha_temp` ha) to `r max_ac_temp` acres (`r max_ha_temp` ha). 

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

```{r map, fig.height = 5}
# map
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
mapview::mapview(
  las_ctg_dta
  , color = "firebrick"
  , lwd = 2
  , col.regions = c("firebrick")
  , alpha.regions = 0.3
  , layer.name = "UAS Flight Boundaries"
  , label = FALSE
  , legend = T
  , popup = FALSE
) +
mapview::mapview(
  harvests
  , zcol = "treatment_type_grp"
  , col.regions = viridis::turbo(n=length(unique(harvests$treatment_type_grp)))
  , alpha.regions = 0.6
  , layer.name = "Harvests (2021)"
  , legend = T
    , popup = leafpop::popupTable(
        harvests
        , zcol = c(
          "year_id"
          , "treatment_type_grp"
          , "activity_name"
        )
        , row.numbers = FALSE
        , feature.id = FALSE
      )
)
```

## Digital Terrain Model (DTM) Map

Map the SfM-derived digital terrain model (DTM) (sometimes called Digital Elevation Models [DEM]) which is a topographic model of the bare Earth containing the elevation data of the terrain.

```{r dtm, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
# aggregate raster and map
    mapview::mapview(
      dtm_rast %>%
        terra::aggregate(fact=2) %>% 
        `*`(3.28) %>% # transform to feet
        stars::st_as_stars()
      , layer.name = "elev. (ft)"
      , col.regions = viridis::viridis(n=100)
      , alpha.regions = 0.7
      , na.color = "transparent"
    )
```

## Canopy height model (CHM) Map

Map the SfM-derived canopy height model which is a measurement of the height of trees above the ground topography.

```{r chm-temp, results='hide', include=FALSE, eval=TRUE}
chm_rast_temp = chm_rast %>%
        terra::aggregate(fact=4) %>% 
        `*`(3.28) %>% # transform to feet
        stars::st_as_stars()
```

```{r chm, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
# aggregate raster and map
    mapview::mapview(
      chm_rast_temp
      , layer.name = "canopy ht. (ft)"
      , col.regions = viridis::plasma(n=50)
      , alpha.regions = 0.7
      , na.color = "transparent"
    ) 
  # # tree_tops
  # mapview::mapview(
  #   treetops_sf_with_dbh
  #   , cex = 0.5
  #   , layer.name = "tree tops"
  #   , label = F
  #   , legend = F
  #   , popup = F
  # )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## DBH Distribution

DBH distribution of trees that are in a harvested unit. Trees within harvest unit boundaries have a DBH range from `r min(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r min(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm) to `r max(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r max(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm) with a median of `r median(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r median(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm). 

```{r dbh-dist}
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(treeID, dbh_in) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = dbh_in)
  ) +
    geom_density(alpha = 0.8, fill = "navy", color = NA) + 
    labs(
      x = "DBH (in)"
      , y = "density"
      , title = "SfM-derived tree DBH distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

## Height Distribution

Height distribution of trees that are in a harvested unit. Trees within harvest unit boundaries have a height range from `r min(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r min(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m) to `r max(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r max(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m) with a median of `r median(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r median(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m). 

```{r ht-dist}
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(treeID, tree_height_ft) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = tree_height_ft)
  ) +
    geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
    labs(
      x = "Height (ft)"
      , y = "density"
      , title = "SfM-derived tree height distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

## Relationship between height and DBH

```{r ht-dbh}
### plot
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::slice_sample(prop = 0.2) %>% 
  dplyr::select(treeID, tree_height_ft, dbh_in) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(y=tree_height_ft, x = dbh_in)
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.6
    , color = "gray"
  ) + 
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , color = "gray33"
    , alpha = 0.7
  ) +
  labs(
    x = "DBH (in)"
    , y = "Tree Ht. (ft)"
    , title = "SfM-derived tree height and DBH relationship"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

# Silvicultural Metrics 

Silvicultural metrics are calculated based on the SfM-derived DBH and height measurements.

```{r silv-calc}
# Common silvicultural metrics are calculated for the entire extent.
  # Note, that stand-level summaries can be computed if stand vector data is provided.
  # metrics include:
    # "n_trees"
    # "stand_area_ha"
    # "trees_per_ha"
    # "mean_dbh_cm"
    # "qmd_cm"
    # "mean_tree_height_m"
    # "loreys_height_m"
    # "basal_area_m2"
    # "basal_area_m2_per_ha"

### stand-level summaries
  silv_metrics = harvests_trees %>%
    sf::st_drop_geometry() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(suid,stand_area_ha) %>%
    dplyr::summarise(
      n_trees = dplyr::n_distinct(treeID)
      , mean_dbh_cm = mean(dbh_cm, na.rm = T)
      , mean_tree_height_m = mean(tree_height_m, na.rm = T)
      , loreys_height_m = sum(basal_area_m2*tree_height_m, na.rm = T) / sum(basal_area_m2, na.rm = T)
      , basal_area_m2 = sum(basal_area_m2, na.rm = T)
      , sum_dbh_cm_sq = sum(dbh_cm^2, na.rm = T)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      trees_per_ha = (n_trees/stand_area_ha)
      , basal_area_m2_per_ha = (basal_area_m2/stand_area_ha)
      , qmd_cm = sqrt(sum_dbh_cm_sq/n_trees)
    ) %>%
    dplyr::select(-c(sum_dbh_cm_sq)) %>%
    # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.28
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    ) %>%
    dplyr::select(
      "suid"
      , "n_trees"
      , "stand_area_ha"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2"
      , "basal_area_m2_per_ha"
      # imperial
      , "stand_area_ac"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2"
      , "basal_area_ft2_per_ac"
    )

### export tabular
  write.csv(
      silv_metrics
      , paste0(delivery_dir, "/stand_silv_metrics.csv")
      , row.names = F
    )

  # join with spatial data
  silv_metrics = harvests %>% 
    # join with plot data data
    dplyr::inner_join(
      silv_metrics
      , by = dplyr::join_by("suid")
    )
```

## Distribution of silvicultural metrics

Distribution of silvicultural metrics across the `r nrow(silv_metrics)` stands in this analysis.

```{r site-metrics}
silv_metrics %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    "suid"
    , "stand_area_ac"
    , "n_trees"
    , "trees_per_ac"
    , "mean_dbh_in"
    , "qmd_in"
    , "mean_tree_height_ft"
    , "loreys_height_ft"
    , "basal_area_ft2_per_ac"
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(suid), names_to = "metric", values_to = "val"
  ) %>% 
  # order metrics
  dplyr::mutate(
    metric = factor(
      metric
      , ordered = T
      , levels = c(
          "stand_area_ac"
          , "n_trees"
          , "trees_per_ac"
          , "mean_tree_height_ft"
          , "loreys_height_ft"
          , "mean_dbh_in"
          , "qmd_in"
          , "basal_area_ft2_per_ac"
        )
      , labels = c(
          latex2exp::TeX("Stand Area (ac)", output = "character")
          , latex2exp::TeX("Number of Trees", output = "character")
          , latex2exp::TeX("Trees $ac^{-1}$")
          , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
          , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
          , latex2exp::TeX("Mean DBH (in)", output = "character")
          , latex2exp::TeX("QMD (in)", output = "character")
          , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
        )
    )
  ) %>% 
  ggplot(
    mapping = aes(
      x = val
      , y = 1
    )
  ) +
  geom_boxplot(fill = "gray", width = 0.7, outlier.shape = NA) +
  geom_point(mapping = aes(group = suid)) +
  facet_wrap(facets = vars(metric), ncol = 2, scales = "free_x", labeller = label_parsed) +
  scale_x_continuous(breaks = scales::extended_breaks(6)) +
  scale_y_continuous(limits = c(0.5,1.5)) +
  labs(
    y = ""
    , x = ""
    , title = "Distribution of Silvicultural Metrics"
    , subtitle = paste(nrow(silv_metrics), "stands included in analysis")
  ) +
  theme_light() + 
  theme(
    legend.position = "none"
    , strip.text = element_text(color = "black")
    , strip.background = element_rect(fill = "gray88")
    , axis.text.y = element_blank()
    , axis.ticks.y = element_blank()
    , panel.grid.major.y = element_blank()
    , panel.grid.minor.y = element_blank()
  )
```

## Stand Basal Area Map

```{r ba-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "basal_area_ft2_per_ac"
      , layer.name = "BA (ft2 per ac)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::cividis(n=nrow(silv_metrics)*1.1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

## Stand QMD Map

Quadratic mean diameter (QMD) is a measure of the diameter of the tree of mean basal area:

$$
\textrm{quadratic mean diameter (QMD)} = \sqrt{\frac{\sum{d_{i}^{2}}}{n}}
$$

, where $d_{i}$ is the diameter at breast height of an individual tree, and $n$ is the total number of trees. 

```{r qmd-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "qmd_in"
      , layer.name = "QMD (in)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::mako(n=nrow(silv_metrics)*1.1, direction = -1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

# Harvested Stand Summaries

Create plots to combine for report

1) orthomosaic + chm + tree tops + stand plot

```{r ortho-plt-fn}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
ortho_plt_fn = function(my_suid){

# convert to stars
  ortho_st = ortho_rast %>%  
    terra::subset(subset = c(1,2,3)) %>%
    terra::crop(
      # stand %>% 
      harvests %>% dplyr::filter(suid==my_suid) %>% 
        sf::st_buffer(20) %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        terra::vect()
    ) %>% 
    terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
    stars::st_as_stars()
  
  # convert to rgb
  ortho_rgb <- stars::st_rgb(
    ortho_st[,,,1:3]
    , dimension = 3
    , use_alpha = FALSE
    # , stretch = "histogram"
    , probs = c(0.005, 0.995)
    , stretch = "percent"
  )
  # ggplot
  plt_rgb <- ggplot() +
    stars::geom_stars(data = ortho_rgb[]) +
    scale_fill_identity(na.value = "transparent") + # !!! don't take this out or RGB plot will kill your computer
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(
      x = ""
      , y = ""
    ) +
    theme_void()
  
  # return(plt_rgb)
  # combine all plot elements
  plt_combine = plt_rgb +
    ggnewscale::new_scale_fill() +
    geom_tile(
      data = chm_rast %>% 
        terra::crop(
          harvests %>% dplyr::filter(suid==my_suid) %>% 
          terra::vect()
        ) %>% 
        terra::mask(
          harvests %>% dplyr::filter(suid==my_suid) %>% 
          terra::vect()
        ) %>% 
        `*`(3.28) %>% # transform to feet
        as.data.frame(xy=T) %>% 
        rename(f=3)
      , mapping = aes(x=x,y=y,fill=f)
    ) +
    scale_fill_viridis_c(option="plasma", alpha = 0.6) +
    geom_sf(
      data = harvests %>% dplyr::filter(suid==my_suid)
      , alpha = 0
      , lwd = 1.5
      , color = "#b22222"
    ) +
    geom_sf(
      data = treetops_sf_with_dbh %>% 
        sf::st_intersection(harvests %>% dplyr::filter(suid==my_suid))
      , color = "#330099"
      , shape = "."
    ) +
    labs(
      fill = "canopy ht. (ft)"
      # , title = "Trees identified within stand boundary"
      , title = "<span><span style='color:#330099;'><b><i>Trees</i></b></span> identified within <span style='color:#b22222;'><b><i>stand</i></b></span> boundary</span>"
    ) +
    theme(
      legend.position = "top" # c(0.5,1)
      , legend.direction = "horizontal"
      , legend.margin = margin(0,0,-1,0)
      , legend.text = element_text(size = 8)
      , legend.title = element_text(size = 8)
      , plot.title = ggtext::element_markdown(size = 10, hjust = 0.5)
    )
  return(plt_combine)
}
# ortho_plt_fn(harvests$suid[2])
```

```{r, include=FALSE, eval=F}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
  xxxxxortho_plt_fn = function(stand,ortho,points,chm){
      # plot ortho
      ortho %>% 
        terra::crop(
          stand %>% 
            sf::st_buffer(20) %>% 
            sf::st_bbox() %>% 
            sf::st_as_sfc() %>% 
            terra::vect()
        ) %>% 
        terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
        terra::plotRGB(
          r = 1, g = 2, b = 3, stretch = "lin", colNA = "transparent"
          , mar = c(4,0,2,0)
          , main = "Trees (blue) identified within stand boundary (red)"
          , cex.main = 0.8
        )
      # plot chm
      chm %>% 
        `*`(3.28) %>% # transform to feet
        terra::crop(stand %>% terra::vect()) %>% 
        terra::mask(stand %>% terra::vect()) %>%
        terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
        terra::plot(
          col = viridis::plasma(50), colNA = NA, alpha = 0.6, add = T
          , plg = list(
            title = "canopy ht. (ft)"
            , title.cex = 0.8
            , title.adj = 0
          )
        )
      # legend(legend = 2,title = "canopy ht. (ft)")
      
      # plot stand
        terra::plot(
          stand %>% terra::vect()
          , col = NA, border = "firebrick", lwd = 2.5
          , add = T
        )
      # plot points
      terra::plot(
        terra::crop(points %>% terra::vect(), stand %>% terra::vect())
        , pch = "."
        , col = "blue"
        , add = T
      )
      
  }
# ortho_1 = function(){ortho_plt_fn(
#     stand = harvests[1,]
#     , ortho = ortho_rast
#     , points = treetops_sf_with_dbh
#     , chm = chm_rast
#   )}
# ortho_1()
```

2) dbh distribution plot

```{r dbh-dist-fn}
dbh_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, dbh_in) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = dbh_in)
    ) +
      geom_density(alpha = 0.8, fill = "navy", color = NA) + 
      labs(
        x = "DBH (in)"
        , y = "density"
        , title = "SfM-derived tree DBH distribution"
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# dbh_dist_fn(harvests$suid[1])
```

3) height distribution plot

```{r ht-dist-fn}
ht_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, tree_height_ft) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = tree_height_ft)
    ) +
      geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
      labs(
        x = "Height (ft)"
        , y = "density"
        , title = "SfM-derived tree height distribution"
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# ht_dist_fn(harvests$suid[1])
```

4) silvicultural metrics imperial table

```{r tbl-fn}
tab_imp_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ac"
      , "n_trees"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2_per_ac"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ac"
            , "n_trees"
            , "trees_per_ac"
            , "mean_tree_height_ft"
            , "loreys_height_ft"
            , "mean_dbh_in"
            , "qmd_in"
            , "basal_area_ft2_per_ac"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ac)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ac^{-1}$")
            , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
            , latex2exp::TeX("Mean DBH (in)", output = "character")
            , latex2exp::TeX("QMD (in)", output = "character")
            , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = "Silvicultural Metrics (imperial)", face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}

# tab_imp_fn(harvests$suid[2])

```

5) silvicultural metrics metric table

```{r tbl-fn2}
tab_met_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ha"
      , "n_trees"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2_per_ha"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ha"
            , "n_trees"
            , "trees_per_ha"
            , "mean_tree_height_m"
            , "loreys_height_m"
            , "mean_dbh_cm"
            , "qmd_cm"
            , "basal_area_m2_per_ha"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ha)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ha^{-1}$")
            , latex2exp::TeX("Mean Tree Height (m)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (m)", output = "character")
            , latex2exp::TeX("Mean DBH (cm)", output = "character")
            , latex2exp::TeX("QMD (cm)", output = "character")
            , latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(
          parse = T, fill = "white", linecolor = "gray"
          , size = 10
        )
      )
    ) %>% 
    ggpubr::tab_add_title(
      text = "Silvicultural Metrics (metric)", face = "bold"
      , size = 10, padding = unit(0.5, "line")
    )
}

# tab_met_fn(harvests$suid[2])
```

6) inset map plot

```{r inset-map}
plt_inset_map_fn <- function(my_suid){
  (
    ggplot() +
      geom_sf(data = bhef_boundary, alpha = 0, lwd = 1, color = "black") +
      geom_sf(data = harvests, alpha = 0, lwd = 0.4, color = "gray") +
      geom_sf(
        data = harvests %>% dplyr::filter(suid == my_suid)
        , fill = "firebrick"
        , alpha = 0.3
        , lwd = 1.5
        , color = "firebrick"
      ) +
      # geom_sf_label(
      #   data = rx_fire
      #   , aes(label = unit)
      #   , label.size = NA
      #   , alpha = 0
      #   , size = 2.5
      # ) +
      xlab("") +
      ylab("") +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_text(size = 5.5)
        , axis.text.x = element_text(angle = 35)
        , panel.border = element_blank()
      )

  )
}
# plt_inset_map_fn(harvests$suid[4])
```

7) contour + elevation map plot

```{r elev-plt-inset}
# sequence by 5's to get 5m breaks list for contour
contour_int_m <- 5
seq_brk <- seq(0, 40000, contour_int_m)
# create contour data
elev_contour <- dtm_rast %>% 
  stars::st_as_stars() %>% 
  stars::st_contour(
    na.rm = TRUE
    , contour_lines = FALSE
    , breaks = # classInt::classIntervals(na.omit(as.vector(x[[1]])))$brks
        seq_brk[
          seq_brk > min(na.omit(as.vector(dtm_rast[[1]]))) - contour_int_m
          & seq_brk < max(na.omit(as.vector(dtm_rast[[1]]))) + contour_int_m
        ]
  )

plt_contour_map_fn <- function(my_suid){
  (
    #########################################
    #########################################
    # map of unit w/ contour
    #########################################
    #########################################
    # plot
    ggplot() +
      geom_tile(
        data = dtm_rast %>% 
          terra::crop(
            harvests %>% dplyr::filter(suid == my_suid) %>% 
              sf::st_buffer(dist = 20) %>% 
              sf::st_bbox() %>% 
              sf::st_as_sfc() %>% 
              terra::vect()
          ) %>% 
          terra::aggregate(fact=2) %>% 
          `*`(3.28) %>% # transform to feet
          as.data.frame(xy=T) %>% 
          dplyr::rename(elev = 3)
        , mapping = aes(x=x,y=y,fill=elev)
      ) +
      geom_sf(data = elev_contour %>% 
        sf::st_crop(
          harvests %>% dplyr::filter(suid == my_suid) %>% 
          sf::st_buffer(dist = 20) %>% 
          sf::st_bbox()
        )
        , alpha = 0
        , lwd = 0.5
        , color = "gray30"
        , show.legend = FALSE
      ) +
      geom_sf(data = harvests %>% dplyr::filter(suid == my_suid)
          , alpha = 0
          , lwd = 1.5
          , color = "firebrick"
          , show.legend = FALSE
      ) +
      scale_fill_viridis_c(
        option = "viridis"
        , alpha = 0.9
        , na.value = "transparent"
        , breaks = scales::extended_breaks(n = 6)
        , labels = scales::comma
      ) +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      labs(
        subtitle = paste0(
          as.character(round(contour_int_m * 3.28, 1))
          , " ft ("
          , as.character(contour_int_m)
          , " m) contours"
        )
        , fill = "elev. (ft)"
        , x = ""
        , y = ""
      ) +
      theme_void() +
      theme(
        legend.position = "top"
        , legend.direction = "horizontal"
        , legend.margin = margin(0,0,-1,0)
        # , legend.box.margin = margin(-1,0,2,0)
        , legend.text = element_text(size = 7, angle = 25, hjust = 0.7)
        , legend.title = element_text(size = 7)
        , plot.subtitle = element_text(
            size=9, face="italic", color="gray20", margin = margin(0,0,0,0)
            , hjust = 0.5
        )
      )
  )
}
# plt_contour_map_fn(harvests$suid[1])
```

build report

```{r report-fn}
report_fn = function(my_suid){
  # generate plots
  ortho_plt_temp = ortho_plt_fn(my_suid) # ortho_plt_fn(harvests$suid[4])
  dbh_dist_temp =  dbh_dist_fn(my_suid) # dbh_dist_fn(harvests$suid[4])
  ht_dist_temp = ht_dist_fn(my_suid) # ht_dist_fn(harvests$suid[4]) 
  plt_inset_map_temp = plt_inset_map_fn(my_suid) # plt_inset_map_fn(harvests$suid[4])
  plt_contour_map_temp = plt_contour_map_fn(my_suid) # plt_contour_map_fn(harvests$suid[4])
  tab_imp_temp = tab_imp_fn(my_suid) # tab_imp_fn(harvests$suid[4])
  tab_met_temp = tab_met_fn(my_suid) # tab_met_fn(harvests$suid[4]) 
  # plot layout
  # area(t,l,b,r)
  layout_temp = c(
    # map
    area(1, 1, 3, 2)
    # space
    , area(1, 3, 1, 4)
    # distributions 
    , area(2, 3, 2, 4)
    , area(3, 3, 3, 4)
    # space
    , area(4, 3, 4, 4)
    # map inset
    , area(4, 1, 5, 1)
    # contour
    , area(4, 2, 5, 2)
    # tables
    , area(5, 3)
    , area(5, 4)
    # space
    , area(6, 1, 6, 4)
  )
  # check the layout
  # plot(layout_temp)
  
  # patchwork 
  rpt = ortho_plt_temp +
    patchwork::plot_spacer() +
    dbh_dist_temp + ht_dist_temp +
    patchwork::plot_spacer() +
    plt_inset_map_temp + 
    plt_contour_map_temp + 
    tab_imp_temp + tab_met_temp +
    patchwork::plot_spacer() +
    patchwork::plot_layout(
      design = layout_temp
      , heights = c(0.01,1,1,0.4,1,0.1)
      , widths = c(0.5,1,1,1)
    )
  # export
  ggplot2::ggsave(
    filename = paste0(delivery_dir, "/suid_", my_suid, ".pdf")
    , plot = rpt
    , device = "pdf"
    , width = 11, height = 8.5, units = "in"
  )
  return(rpt)
}
```

```{r print-reports, results='hide', fig.show='asis', fig.height=8}
# generate and print reports
harvests$suid %>% 
  purrr::map(report_fn) %>% 
  print()
```

```{r pdf-comb, results='hide'}
# combine all pdfs together
# combine with pdftools
pdftools::pdf_combine(
  sort(list.files(delivery_dir, pattern = "\\.pdf$", full.names = TRUE))
  , output = paste0(
    delivery_dir
    , "/BHEF_UAS_202306_report_"
    , gsub("-", "", Sys.Date())
    , ".pdf"
  )
)
```

```{r include=FALSE, eval=FALSE}
# Field Validation Data

Here are both the overstory and regen data from the Voodoo stands. 
Each tree has UTM coordinates and then each regen subplot has UTM coordinates. 
Also attached is my script for matching UAS and stem-mapped trees. 
I have only used the script when comparing complete stem-maps with UAS and not with plots, likely to be something we will need to think through to make it work.

```


```{r include=FALSE, eval=FALSE}
library(patchwork)
layout_temp = c(
    # map
    area(1, 1, 3, 2)
    # space
    , area(1, 3, 1, 4)
    # distributions 
    , area(2, 3, 2, 4)
    , area(3, 3, 3, 4)
    # space
    , area(4, 3, 4, 4)
    # map inset
    , area(4, 1, 5, 1)
    # contour
    , area(4, 2, 5, 2)
    # tables
    , area(5, 3)
    , area(5, 4)
    # space
    , area(6, 1, 6, 4)
  )
  # check the layout
  plot(layout_temp)
  
```
