---
title: "Black Hills Experimental Forest UAS Mission Data Summary"
author: "George Woolsey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'index.html')) 
  })
---

# Introduction

This project summarizes processed point cloud data created using structure from motion (SfM) photogrammetry methods on imagery collected from Unmanned Aircraft System (UAS) flights completed in June 2023. UAS imagery was collected using a DJI Phantom 4 Pro equipped with a 20 megapixel metal oxide semiconductor red-green-blue camera at a fixed 8.8 mm focal length. Flights followed a pre-programmed serpentine flight paths at an altitude of 120 m above ground level, using a nadir camera orientation, with 80% forward and side image overlap.The UAV imagery was processed in Agisoft Metashape using the routine outlined by [Tinkham and Swayze (2021)](https://scholar.google.com/scholar?cluster=11260597505702247290&hl=en&as_sdt=0,6) to produce an SfM point cloud for optimized tree detection. 

The SfM-derived point cloud was processed using the methods outlined by [Swayze and Tinkham (2022)](https://scholar.google.com/scholar?oi=bibs&hl=en&cluster=10655866445299954513). Specifically, [this script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/da5c8a13e9b2f2914b306201605e907643b3006d/src/point_cloud_processing.R) which is outlined in detail at it's [parent Github repository](https://georgewoolsey.github.io/point_cloud_tree_detection_ex/) was used to process the raw point cloud data. UAS point cloud files (`laz` files) were processed in two batches with the outputs combined using [this script](https://github.com/georgewoolsey/point_cloud_tree_detection_ex/blob/6813f499e127b474dd91b480a993beb3d5225a11/src/point_cloud_processing_run_combine.R) from the same repository.

SfM techniques using UAS data reliably identify most tree locations and heights in open-canopied forests. However, their ability to characterize diameter at breast height (DBH) is limited to estimates of a fraction of trees within the area. SfM-derived DBH was estimated at the tree level following the methods outlined by [Tinkham et al. (2022)](https://scholar.google.com/scholar?cluster=14807895789640069059&hl=en&as_sdt=0,6).

## Objective

The objective of this analysis is to summarize forest stand characteristics using the UAS-derived tree list which includes tree location, height, and DBH measurements. Forest stands in this analysis include stands which were harvested in 2021 using single-tree and group-selection silvicultural prescriptions. 

The Forest Activity Tracking System (FACTS) [database](https://data.fs.usda.gov/geodata/edw/datasets.php?xmlKeyword) maintained by the U.S. Department of Agriculture, Forest Service (USFS) was used to delineate georeferenced boundaries of forest harvest activities.

# Setup
  
```{r, include=FALSE, warning=F, message=F}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  # , results='hide'
  , fig.width = 10
  , fig.height = 7
)
```

## Load packages

Load all packages used in program.

```{r pkg-load}
# library
library(tidyverse) # the tidyverse
library(viridis) # viridis colors
library(scales) # work with number and plot scales
library(latex2exp) # math formulas text
# spatial analysis
library(terra) # raster
library(sf) # simple features
library(stars) # stars raster
# visualization
library(kableExtra)
library(patchwork) # ! better align plots in grid
library(mapview) # interactive html maps
library(leafpop) # interactive html maps
library(ggnewscale) # add another scale to ggplot

# option to put satellite imagery as base layer of mapview maps
  mapview::mapviewOptions(
    homebutton = FALSE
    , basemaps = c("Esri.WorldImagery","OpenStreetMap")
  )
```


# Data Load and Setup

```{r dta-load, results='hide'}
# where is the processed data from point_cloud_processing_run_combine.R ?
input_dir = "../data/point_cloud_processing_BHEF_202306_combined"
# set output directory
delivery_dir = "../data/delivery"
dir.create(delivery_dir, showWarnings = FALSE)

### read data from point_cloud_processing
#rasters
dtm_rast = terra::rast(paste0(input_dir, "/dtm_1m.tif"))
chm_rast = terra::rast(paste0(input_dir, "/chm_0.25m.tif"))
# crowns = terra::rast(paste0(input_dir, "/top_down_detected_tree_crowns.tif"))

# las extent
las_ctg_dta = sf::st_read(paste0(input_dir, "/raw_las_ctg_info.gpkg"))

# vectors
# crown polygons
  crowns_sf_with_dbh = sf::st_read(paste0(input_dir, "/final_detected_crowns.gpkg")) %>% 
    dplyr::select(-c(basal_area_ft2)) %>% 
    dplyr::rename(
      reg_est_lower_dbh_cm = reg_est_dbh_cm_lower
      , reg_est_upper_dbh_cm = reg_est_dbh_cm_upper
    ) %>% 
    # convert to imperial units
      dplyr::mutate(
        dplyr::across(
          .cols = tidyselect::ends_with("_cm")
          , ~ .x * 0.394
          , .names = "{.col}_in"
        )
        , dplyr::across(
          .cols = tidyselect::ends_with("_m")
          , ~ .x * 3.28
          , .names = "{.col}_ft"
        )
        , dplyr::across(
          .cols = tidyselect::ends_with("_m2_per_ha")
          , ~ .x * 4.359
          , .names = "{.col}_ftac"
        )
        , dplyr::across(
          .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
          , ~ .x * 0.405
          , .names = "{.col}_ac"
        )
        , dplyr::across(
          .cols = tidyselect::ends_with("_area_ha")
          , ~ .x * 2.471
          , .names = "{.col}_ac"
        )
        , dplyr::across(
          .cols = tidyselect::ends_with("_m2")
          , ~ .x * 10.764
          , .names = "{.col}_ft2"
        )
      ) %>%
      dplyr::rename_with(
        .fn = function(x){dplyr::case_when(
          stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
          , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
          , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
          , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
          , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
          , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
          , TRUE ~ x
        )}
      )

# tree top points
treetops_sf_with_dbh = sf::st_read(paste0(input_dir, "/final_detected_tree_tops.gpkg")) %>% 
  dplyr::select(-c(basal_area_ft2)) %>% 
  dplyr::rename(
    reg_est_lower_dbh_cm = reg_est_dbh_cm_lower
    , reg_est_upper_dbh_cm = reg_est_dbh_cm_upper
  ) %>% 
  # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.28
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    )

# set crs
proj_crs = sf::st_crs(las_ctg_dta)

# read data from FACTS
# harvests
  harvests = sf::st_read("../data/bhef_harvests.gpkg") %>% 
    dplyr::filter(
      year_id == 2021
      # year_id >= year(Sys.time()) - 15 
      & !(treatment_type_grp %in% c("Improvement/Liberation Cut", "Other", "Sanitation Cut"))
    ) %>% 
    sf::st_transform(proj_crs)
# bhef bounds
  bhef_boundary = sf::st_read("../data/bhef_boundary.gpkg") %>% 
    sf::st_transform(proj_crs)
```

## Load orthomosaic rasters

```{r ortho-ld, results='hide'}
# read list of orthos
ortho_list_temp = list.files("../data/ortho/", pattern = ".*\\.(tif|tiff)$", full.names = T)[] %>% 
    purrr::map(function(x){terra::rast(x)})

# ortho_list_temp[[1]] %>%
#   terra::aggregate(20) %>%
#   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "hist", colNA = "transparent")

####### ensure the resolution of the rasters matches 
  # terra::res(ortho_list_temp[[1]])
  
  ## function
  change_res_fn = function(r, my_res=1){
    r2 = r
    terra::res(r2) = my_res
    r2 = terra::resample(r, r2)
    return(r2)
  }
  ## apply the function
  ortho_list_temp = 1:length(ortho_list_temp) %>% 
    purrr::map(function(x){change_res_fn(ortho_list_temp[[x]], my_res=0.5)})
  
  # terra::res(ortho_list_temp[[1]])
  
  # ortho_list_temp[[1]] %>%
  #   terra::aggregate(2) %>%
  #   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "hist", colNA = "transparent")

######## mosaic the raster list 
  ortho_rast = terra::mosaic(
    terra::sprc(ortho_list_temp)
    , fun = "min" # min only thing that works
  ) 

  # ortho_rast %>%
  #   # terra::aggregate(2) %>%
  #   terra::plotRGB(r = 1, g = 2, b = 3, stretch = "lin", colNA = "transparent")
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## Combine harvest units with tree locations

Spatially combine harvest units with tree locations. One tree can be located in multiple harvest units if the harvest units spatially overlap; I'll allow it. Trees are matched to a harvest unit based on the location of the tree top which may or may not align with the location of the tree bole at DBH.

```{r harvest-trees}
#################################################################################
#################################################################################
# Join tree tops with forest stands
#################################################################################
#################################################################################
harvests_trees = harvests %>%
  dplyr::mutate(
    stand_area_m2 = sf::st_area(.) %>% as.numeric()
    , stand_area_ha = stand_area_m2/10000
  ) %>% 
  sf::st_intersection(las_ctg_dta) %>% 
  dplyr::mutate(
    intrsct_stand_area_m2 = sf::st_area(.) %>% as.numeric()
  ) %>% 
  dplyr::filter(round(intrsct_stand_area_m2, 0) == round(stand_area_m2, 0)) %>% 
  dplyr::select(
    suid, forest_commonname, admin_region_code, activity_name
    , treatment_type, treatment_type_grp, date_compl, year_id
    , stand_area_m2, stand_area_ha
  ) %>% 
  sf::st_intersection(treetops_sf_with_dbh)
# filter harvests
harvests = harvests %>% 
  dplyr::filter(suid %in% unique(harvests_trees$suid))

```

# Exploratory Analysis

High-level summaries of the data.

## Study Location Map
```{r stats-temp, include=FALSE, eval=TRUE}
ac_temp = las_ctg_dta %>% sf::st_area() %>% as.numeric() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
ha_temp = las_ctg_dta %>% sf::st_area() %>% as.numeric() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
min_ac_temp = harvests %>% sf::st_area() %>% as.numeric() %>% min() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
min_ha_temp = harvests %>% sf::st_area() %>% as.numeric() %>% min() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
max_ac_temp = harvests %>% sf::st_area() %>% as.numeric() %>% max() %>% `/`(4046.856) %>% scales::comma(accuracy = 0.1)
max_ha_temp = harvests %>% sf::st_area() %>% as.numeric() %>% max() %>% `/`(10000) %>% scales::comma(accuracy = 0.1)
```

Map of the BHEF, 2021 harvest boundaries, and 2023-06 UAS flight boundaries. The UAS flights covered `r ac_temp` acres (`r ha_temp` ha). There are `r nrow(harvests)` stands harvested in 2021 that are entirely within the UAS flight boundaries. The harvested stands range in size from `r min_ac_temp` acres (`r min_ha_temp` ha) to `r max_ac_temp` acres (`r max_ha_temp` ha). 

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

```{r map, fig.height = 5}
# map
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
mapview::mapview(
  las_ctg_dta
  , color = "firebrick"
  , lwd = 2
  , col.regions = c("firebrick")
  , alpha.regions = 0.3
  , layer.name = "UAS Flight Boundaries"
  , label = FALSE
  , legend = T
  , popup = FALSE
) +
mapview::mapview(
  harvests
  , zcol = "treatment_type_grp"
  , col.regions = viridis::turbo(n=length(unique(harvests$treatment_type_grp)))
  , alpha.regions = 0.6
  , layer.name = "Harvests (2021)"
  , legend = T
    , popup = leafpop::popupTable(
        harvests
        , zcol = c(
          "year_id"
          , "treatment_type_grp"
          , "activity_name"
        )
        , row.numbers = FALSE
        , feature.id = FALSE
      )
)
```

## Digital Terrain Model (DTM) Map

Map the SfM-derived digital terrain model (DTM) (sometimes called Digital Elevation Models [DEM]) which is a topographic model of the bare Earth containing the elevation data of the terrain.

```{r dtm, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
# aggregate raster and map
    mapview::mapview(
      dtm_rast %>%
        terra::aggregate(fact=2) %>% 
        `*`(3.28) %>% # transform to feet
        stars::st_as_stars()
      , layer.name = "elev. (ft)"
      , col.regions = viridis::viridis(n=100)
      , alpha.regions = 0.7
      , na.color = "transparent"
    )
```

## Canopy height model (CHM) Map

Map the SfM-derived canopy height model which is a measurement of the height of trees above the ground topography.

```{r chm-temp, results='hide', include=FALSE, eval=TRUE}
chm_rast_temp = chm_rast %>%
        terra::aggregate(fact=4) %>% 
        `*`(3.28) %>% # transform to feet
        stars::st_as_stars()
```

```{r chm, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
# aggregate raster and map
    mapview::mapview(
      chm_rast_temp
      , layer.name = "canopy ht. (ft)"
      , col.regions = viridis::plasma(n=50)
      , alpha.regions = 0.7
      , na.color = "transparent"
    ) 
  # # tree_tops
  # mapview::mapview(
  #   treetops_sf_with_dbh
  #   , cex = 0.5
  #   , layer.name = "tree tops"
  #   , label = F
  #   , legend = F
  #   , popup = F
  # )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

## DBH Distribution

DBH distribution of trees that are in a harvested unit. Trees within harvest unit boundaries have a DBH range from `r min(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r min(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm) to `r max(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r max(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm) with a median of `r median(harvests_trees$dbh_in) %>% scales::comma(accuracy = 0.1)` in (`r median(harvests_trees$dbh_cm) %>% scales::comma(accuracy = 0.1)` cm). 

```{r dbh-dist}
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(treeID, dbh_in) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = dbh_in)
  ) +
    geom_density(alpha = 0.8, fill = "navy", color = NA) + 
    labs(
      x = "DBH (in)"
      , y = "density"
      , title = "SfM derived tree DBH distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

## Height Distribution

Height distribution of trees that are in a harvested unit. Trees within harvest unit boundaries have a height range from `r min(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r min(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m) to `r max(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r max(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m) with a median of `r median(harvests_trees$tree_height_ft) %>% scales::comma(accuracy = 0.1)` ft (`r median(harvests_trees$tree_height_m) %>% scales::comma(accuracy = 0.1)` m). 

```{r ht-dist}
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(treeID, tree_height_ft) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(x = tree_height_ft)
  ) +
    geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
    labs(
      x = "Height (ft)"
      , y = "density"
      , title = "SfM derived tree height distribution"
    ) +
    scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
    theme_light() +
    theme(
      legend.position = "none"
    )
```

## Relationship between height and DBH

```{r ht-dbh}
### plot
harvests_trees %>% 
  sf::st_drop_geometry() %>% 
  dplyr::slice_sample(prop = 0.5) %>% 
  dplyr::select(treeID, tree_height_ft, dbh_in) %>% 
  dplyr::distinct() %>% 
  ggplot(
    mapping = aes(y=tree_height_ft, x = dbh_in)
  ) +
  geom_point(
    alpha = 0.6
    , size = 0.5
    , color = "gray"
  ) + 
  geom_smooth(
    method = "loess"
    , se = F
    , span = 1
    , color = "gray33"
    , alpha = 0.7
  ) +
  labs(
    x = "DBH (in)"
    , y = "Tree Ht. (ft)"
    , title = "SfM derived tree height and DBH relationship"
  ) +
  theme_light() +
  theme(
    legend.position = "none"
  )
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("_temp",ls())])
gc()
```

# Silvicultural Metrics 

Silvicultural metrics are calculated based on the SfM-derived DBH and height measurements.

```{r silv-calc}
# Common silvicultural metrics are calculated for the entire extent.
  # Note, that stand-level summaries can be computed if stand vector data is provided.
  # metrics include:
    # "n_trees"
    # "stand_area_ha"
    # "trees_per_ha"
    # "mean_dbh_cm"
    # "qmd_cm"
    # "mean_tree_height_m"
    # "loreys_height_m"
    # "basal_area_m2"
    # "basal_area_m2_per_ha"

### stand-level summaries
  silv_metrics = harvests_trees %>%
    sf::st_drop_geometry() %>%
    dplyr::ungroup() %>%
    dplyr::group_by(suid,stand_area_ha) %>%
    dplyr::summarise(
      n_trees = dplyr::n_distinct(treeID)
      , mean_dbh_cm = mean(dbh_cm, na.rm = T)
      , mean_tree_height_m = mean(tree_height_m, na.rm = T)
      , loreys_height_m = sum(basal_area_m2*tree_height_m, na.rm = T) / sum(basal_area_m2, na.rm = T)
      , basal_area_m2 = sum(basal_area_m2, na.rm = T)
      , sum_dbh_cm_sq = sum(dbh_cm^2, na.rm = T)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      trees_per_ha = (n_trees/stand_area_ha)
      , basal_area_m2_per_ha = (basal_area_m2/stand_area_ha)
      , qmd_cm = sqrt(sum_dbh_cm_sq/n_trees)
    ) %>%
    dplyr::select(-c(sum_dbh_cm_sq)) %>%
    # convert to imperial units
    dplyr::mutate(
      dplyr::across(
        .cols = tidyselect::ends_with("_cm")
        , ~ .x * 0.394
        , .names = "{.col}_in"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m")
        , ~ .x * 3.28
        , .names = "{.col}_ft"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 4.359
        , .names = "{.col}_ftac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_per_ha") & !tidyselect::ends_with("_m2_per_ha")
        , ~ .x * 0.405
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_area_ha")
        , ~ .x * 2.471
        , .names = "{.col}_ac"
      )
      , dplyr::across(
        .cols = tidyselect::ends_with("_m2")
        , ~ .x * 10.764
        , .names = "{.col}_ft2"
      )
    ) %>%
    dplyr::rename_with(
      .fn = function(x){dplyr::case_when(
        stringr::str_ends(x,"_cm_in") ~ stringr::str_replace(x,"_cm_in","_in")
        , stringr::str_ends(x,"_m_ft") ~ stringr::str_replace(x,"_m_ft","_ft")
        , stringr::str_ends(x,"_m2_per_ha_ftac") ~ stringr::str_replace(x,"_m2_per_ha_ftac","_ft2_per_ac")
        , stringr::str_ends(x,"_per_ha_ac") ~ stringr::str_replace(x,"_per_ha_ac","_per_ac")
        , stringr::str_ends(x,"_area_ha_ac") ~ stringr::str_replace(x,"_area_ha_ac","_area_ac")
        , stringr::str_ends(x,"_m2_ft2") ~ stringr::str_replace(x,"_m2_ft2","_ft2")
        , TRUE ~ x
      )}
    ) %>%
    dplyr::select(
      "suid"
      , "n_trees"
      , "stand_area_ha"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2"
      , "basal_area_m2_per_ha"
      # imperial
      , "stand_area_ac"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2"
      , "basal_area_ft2_per_ac"
    )

### export tabular
  write.csv(
      silv_metrics
      , paste0(delivery_dir, "/stand_silv_metrics.csv")
      , row.names = F
    )

  # join with spatial data
  silv_metrics = harvests %>% 
    # join with plot data data
    dplyr::inner_join(
      silv_metrics
      , by = dplyr::join_by("suid")
    )
```

## Distribution of silvicultural metrics

Distribution of silvicultural metrics across the `r nrow(silv_metrics)` stands in this analysis.

```{r site-metrics}
silv_metrics %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(
    "suid"
    , "stand_area_ac"
    , "n_trees"
    , "trees_per_ac"
    , "mean_dbh_in"
    , "qmd_in"
    , "mean_tree_height_ft"
    , "loreys_height_ft"
    , "basal_area_ft2_per_ac"
  ) %>% 
  tidyr::pivot_longer(
    cols = -c(suid), names_to = "metric", values_to = "val"
  ) %>% 
  # order metrics
  dplyr::mutate(
    metric = factor(
      metric
      , ordered = T
      , levels = c(
          "stand_area_ac"
          , "n_trees"
          , "trees_per_ac"
          , "mean_tree_height_ft"
          , "loreys_height_ft"
          , "mean_dbh_in"
          , "qmd_in"
          , "basal_area_ft2_per_ac"
        )
      , labels = c(
          latex2exp::TeX("Stand Area (ac)", output = "character")
          , latex2exp::TeX("Number of Trees", output = "character")
          , latex2exp::TeX("Trees $ac^{-1}$")
          , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
          , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
          , latex2exp::TeX("Mean DBH (in)", output = "character")
          , latex2exp::TeX("QMD (in)", output = "character")
          , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
        )
    )
  ) %>% 
  ggplot(
    mapping = aes(
      x = val
      , y = 1
    )
  ) +
  geom_boxplot(fill = "gray", width = 0.7, outlier.shape = NA) +
  geom_point(mapping = aes(group = suid)) +
  facet_wrap(facets = vars(metric), ncol = 2, scales = "free_x", labeller = label_parsed) +
  scale_x_continuous(breaks = scales::extended_breaks(6)) +
  scale_y_continuous(limits = c(0.5,1.5)) +
  labs(
    y = ""
    , x = ""
    , title = "Distribution of Silvicultural Metrics"
    , subtitle = paste(nrow(silv_metrics), "stands included in analysis")
  ) +
  theme_light() + 
  theme(
    legend.position = "none"
    , strip.text = element_text(color = "black")
    , strip.background = element_rect(fill = "gray88")
    , axis.text.y = element_blank()
    , axis.ticks.y = element_blank()
    , panel.grid.major.y = element_blank()
    , panel.grid.minor.y = element_blank()
  )
```

## Stand Basal Area Map

```{r ba-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "basal_area_ft2_per_ac"
      , layer.name = "BA (ft2 per ac)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::cividis(n=nrow(silv_metrics)*1.1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

## Stand QMD Map

Quadratic mean diameter (QMD) is a measure of the diameter of the tree of mean basal area:

$$
\textrm{quadratic mean diameter (QMD)} = \sqrt{\frac{\sum{d_{i}^{2}}}{n}}
$$

, where $d_{i}$ is the diameter at breast height of an individual tree, and $n$ is the total number of trees. 

```{r qmd-map, fig.height = 5}
mapview::mapview(
  bhef_boundary
  , color = "black"
  , lwd = 3
  , alpha.regions = 0
  , layer.name = "BHEF"
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
) +
    mapview::mapview(
      silv_metrics
      , zcol = "qmd_in"
      , layer.name = "QMD (in)" #latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
      , col.regions = viridis::mako(n=nrow(silv_metrics)*1.1, direction = -1)
      , alpha.regions = 0.7
      , na.color = "transparent"
      , popup = F
    ) 
```

# Harvested Stand Summaries

Create plots to combine for report

```{r ortho-plt-fn}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
ortho_plt_fn = function(my_suid){

# convert to stars
  ortho_st = ortho_rast %>%  
    terra::subset(subset = c(1,2,3)) %>%
    terra::crop(
      # stand %>% 
      harvests %>% dplyr::filter(suid==my_suid) %>% 
        sf::st_buffer(20) %>% 
        sf::st_bbox() %>% 
        sf::st_as_sfc() %>% 
        terra::vect()
    ) %>% 
    terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
    stars::st_as_stars()
  
  # convert to rgb
  ortho_rgb <- stars::st_rgb(
    ortho_st[,,,1:3]
    , dimension = 3
    , use_alpha = FALSE
    # , stretch = "histogram"
    , probs = c(0.005, 0.995)
    , stretch = "percent"
  )
  # ggplot
  plt_rgb <- ggplot() +
    stars::geom_stars(data = ortho_rgb[]) +
    scale_fill_identity(na.value = "transparent") + # !!! don't take this out or RGB plot will kill your computer
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    labs(
      x = ""
      , y = ""
    ) +
    theme_void()
  
  # return(plt_rgb)
  # combine all plot elements
  plt_combine = plt_rgb +
    ggnewscale::new_scale_fill() +
    geom_tile(
      data = chm_rast %>% 
        terra::crop(
          harvests %>% dplyr::filter(suid==my_suid) %>% 
          terra::vect()
        ) %>% 
        terra::mask(
          harvests %>% dplyr::filter(suid==my_suid) %>% 
          terra::vect()
        ) %>% 
        `*`(3.28) %>% # transform to feet
        as.data.frame(xy=T) %>% 
        rename(f=3)
      , mapping = aes(x=x,y=y,fill=f)
    ) +
    scale_fill_viridis_c(option="plasma", alpha = 0.6) +
    geom_sf(
      data = harvests %>% dplyr::filter(suid==my_suid)
      , alpha = 0
      , lwd = 2
      , color = "#b22222"
    ) +
    geom_sf(
      data = treetops_sf_with_dbh %>% 
        sf::st_intersection(harvests %>% dplyr::filter(suid==my_suid))
      , color = "#330099"
      , shape = "."
    ) +
    labs(
      fill = "canopy ht. (ft)"
      # , title = "Trees identified within stand boundary"
      , title = "<span><span style='color:#330099;'><b><i>Trees</i></b></span> identified within <span style='color:#b22222;'><b><i>stand</i></b></span> boundary</span>"
    ) +
    theme(
      legend.position = "bottom" # c(0.5,1)
      , legend.direction = "horizontal"
      , legend.text = element_text(size = 8)
      , legend.title = element_text(size = 8)
      , plot.title = ggtext::element_markdown(size = 10)
    )
  return(plt_combine)
}
# ortho_plt_fn(harvests$suid[2])
```

```{r, include=FALSE, eval=F}
######################################################################################
# function to plot ortho + chm + points + stand
######################################################################################
  ortho_plt_fn = function(stand,ortho,points,chm){
      # plot ortho
      ortho %>% 
        terra::crop(
          stand %>% 
            sf::st_buffer(20) %>% 
            sf::st_bbox() %>% 
            sf::st_as_sfc() %>% 
            terra::vect()
        ) %>% 
        terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
        terra::plotRGB(
          r = 1, g = 2, b = 3, stretch = "lin", colNA = "transparent"
          , mar = c(4,0,2,0)
          , main = "Trees (blue) identified within stand boundary (red)"
          , cex.main = 0.8
        )
      # plot chm
      chm %>% 
        `*`(3.28) %>% # transform to feet
        terra::crop(stand %>% terra::vect()) %>% 
        terra::mask(stand %>% terra::vect()) %>%
        terra::aggregate(fact = 2, fun = "mean", na.rm = T) %>% 
        terra::plot(
          col = viridis::plasma(50), colNA = NA, alpha = 0.6, add = T
          , plg = list(
            title = "canopy ht. (ft)"
            , title.cex = 0.8
            , title.adj = 0
          )
        )
      # legend(legend = 2,title = "canopy ht. (ft)")
      
      # plot stand
        terra::plot(
          stand %>% terra::vect()
          , col = NA, border = "firebrick", lwd = 2.5
          , add = T
        )
      # plot points
      terra::plot(
        terra::crop(points %>% terra::vect(), stand %>% terra::vect())
        , pch = "."
        , col = "blue"
        , add = T
      )
      
  }
# ortho_1 = function(){ortho_plt_fn(
#     stand = harvests[1,]
#     , ortho = ortho_rast
#     , points = treetops_sf_with_dbh
#     , chm = chm_rast
#   )}
# ortho_1()
```

```{r dbh-dist-fn}
dbh_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, dbh_in) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = dbh_in)
    ) +
      geom_density(alpha = 0.8, fill = "navy", color = NA) + 
      labs(
        x = "DBH (in)"
        , y = "density"
        , title = "SfM derived tree DBH distribution"
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=20)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# dbh_dist_fn(harvests$suid[1])
```

```{r ht-dist-fn}
ht_dist_fn = function(my_suid){
  harvests_trees %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(treeID, tree_height_ft) %>% 
    dplyr::distinct() %>% 
    ggplot(
      mapping = aes(x = tree_height_ft)
    ) +
      geom_density(alpha = 0.8, fill = "steelblue", color = NA) + 
      labs(
        x = "Height (ft)"
        , y = "density"
        , title = "SfM derived tree height distribution"
      ) +
      scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
      theme_light() +
      theme(
        legend.position = "none"
      )
}
# ht_dist_fn(harvests$suid[1])
```

```{r tbl-fn}
tab_imp_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ac"
      , "n_trees"
      , "trees_per_ac"
      , "mean_dbh_in"
      , "qmd_in"
      , "mean_tree_height_ft"
      , "loreys_height_ft"
      , "basal_area_ft2_per_ac"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ac"
            , "n_trees"
            , "trees_per_ac"
            , "mean_tree_height_ft"
            , "loreys_height_ft"
            , "mean_dbh_in"
            , "qmd_in"
            , "basal_area_ft2_per_ac"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ac)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ac^{-1}$")
            , latex2exp::TeX("Mean Tree Height (ft)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (ft)", output = "character")
            , latex2exp::TeX("Mean DBH (in)", output = "character")
            , latex2exp::TeX("QMD (in)", output = "character")
            , latex2exp::TeX("Basal Area $ft^{2} \\cdot ac^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(parse = T, fill = "white", linecolor = "gray")
      )
    ) %>% 
    ggpubr::tab_add_title(text = "Silvicultural Metrics (imperial)", face = "bold", padding = unit(0.5, "line"))
}

# tab_imp_fn(harvests$suid[2])

```

```{r tbl-fn2}
tab_met_fn = function(my_suid){
  silv_metrics %>% 
    dplyr::filter(suid == my_suid) %>% 
    sf::st_drop_geometry() %>% 
    dplyr::select(
      "suid"
      , "stand_area_ha"
      , "n_trees"
      , "trees_per_ha"
      , "mean_dbh_cm"
      , "qmd_cm"
      , "mean_tree_height_m"
      , "loreys_height_m"
      , "basal_area_m2_per_ha"
    ) %>% 
    tidyr::pivot_longer(
      cols = -c(suid), names_to = "metric", values_to = "val"
    ) %>% 
    # order metrics
    dplyr::mutate(
      metric = factor(
        metric
        , ordered = T
        , levels = c(
            "stand_area_ha"
            , "n_trees"
            , "trees_per_ha"
            , "mean_tree_height_m"
            , "loreys_height_m"
            , "mean_dbh_cm"
            , "qmd_cm"
            , "basal_area_m2_per_ha"
          )
        , labels = c(
            latex2exp::TeX("Stand Area (ha)", output = "character")
            , latex2exp::TeX("Number of Trees", output = "character")
            , latex2exp::TeX("Trees $ha^{-1}$")
            , latex2exp::TeX("Mean Tree Height (m)", output = "character")
            , latex2exp::TeX("Lorey's Mean Height (m)", output = "character")
            , latex2exp::TeX("Mean DBH (cm)", output = "character")
            , latex2exp::TeX("QMD (cm)", output = "character")
            , latex2exp::TeX("Basal Area $m^{2} \\cdot ha^{-1}$")
          )
      )
    ) %>% 
    dplyr::mutate(val = scales::comma(val,accuracy=0.1)) %>% 
    dplyr::select(metric, val) %>% 
    ggpubr::ggtexttable(
      rows = NULL
      , cols = rep("", 2)
      , theme = ggpubr::ttheme(
        base_style = "blank"
        , tbody.style = ggpubr::tbody_style(parse = T, fill = "white", linecolor = "gray")
      )
    ) %>% 
    ggpubr::tab_add_title(text = "Silvicultural Metrics (metric)", face = "bold", padding = unit(0.5, "line"))
}

# tab_met_fn(harvests$suid[2])
```

```{r inset-map}
plt_inset_map_fn <- function(my_suid){
  (
    ggplot() +
      geom_sf(data = bhef_boundary, alpha = 0, lwd = 1, color = "black") +
      geom_sf(
        data = harvests %>% dplyr::filter(suid == my_suid)
        , fill = "firebrick"
        , alpha = 0.3
        , lwd = 1.5
        , color = "firebrick"
      ) +
      # geom_sf_label(
      #   data = rx_fire
      #   , aes(label = unit)
      #   , label.size = NA
      #   , alpha = 0
      #   , size = 2.5
      # ) +
      xlab("") +
      ylab("") +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme_bw() +
      theme(
        axis.text = element_text(size = 5.5)
        , axis.text.x = element_text(angle = 35)
        , panel.border = element_blank()
      )

  )
}
# plt_inset_map_fn(harvests$suid[2])
```


```{r include=FALSE, eval=FALSE}
ortho_plt_temp = ortho_plt_fn(harvests$suid[1])
dbh_dist_temp = dbh_dist_fn(harvests$suid[1])
ht_dist_temp = ht_dist_fn(harvests$suid[1]) 
plt_inset_map_temp = plt_inset_map_fn(harvests$suid[1])
tab_imp_temp = tab_imp_fn(harvests$suid[1])
tab_met_temp = tab_met_fn(harvests$suid[1]) 


  ortho_plt_temp / 
    (plt_inset_map_temp + patchwork::plot_spacer()) + 
      patchwork::plot_layout(ncol = 1, heights = c(1,0.3))
(ortho_plt_temp / patchwork::plot_spacer()) + patchwork::plot_layout(ncol = 1, heights = c(1,0.1)) 

  xxx = (ortho_plt_temp + patchwork::plot_spacer() + patchwork::plot_layout(ncol = 1, heights = c(1,0.1))) 
  
  xxx 
  
    ((tab_imp_temp | tab_met_temp)  /
    (dbh_dist_temp | ht_dist_temp)) & 
    theme(plot.margin = margin(1,1,1,1,"cm"))
  +
    patchwork::plot_layout(widths = c(1.5,1))
   +
    
  (
    (tab_imp_temp | tab_met_temp) /
    (dbh_dist_temp | ht_dist_temp)
  ) & theme(plot.margin = margin(1,1,1,1,"cm"))
  +
  patchwork::plot_layout(nrow = 1, widths = c(1,1.2))
ggplot2::ggsave(paste0("C:/Users/georg/Downloads/hey_xxx.jpeg"),width = 11, height = 8.5, units = "in")

```



```{r include=FALSE, eval=FALSE}
ggpubr::ggtexttable(df, rows = NULL, theme = ttheme("light")) %>% 
  ggpubr::tab_add_title(text = "Silvicultural Metrics (imperial)", face = "bold", padding = unit(0.1, "line"))

## summary table
silv_metrics %>%
  sf::st_drop_geometry() %>%
  dplyr::filter(plotID==silv_metrics_temp$plotID[1]) %>%
  tidyr::pivot_longer(
    cols = -c(plotID)
    , names_to = "measure"
    , values_to = "value"
  ) %>%
  kableExtra::kbl(
    caption = "Silvicultural metrics in both metric and imperial units"
    , digits = 2
  ) %>%
  kableExtra::kable_styling()



```

